### 2018-02-05 write-up/background.tex
Start Dissertation
At some point I should put this under continuous-integration as well?

\chapter{Background}

\begin{guidance}
    A more extensive coverage of what's required to understand your
    work. In general you should assume the reader has a good undergraduate
    degree in computer science, but is not necessarily an expert in
    the particular area you've been working on. Hence this chapter
    may need to summarize some ``text book'' material.

    This is not something you'd normally require in an academic paper,
    and it may not be appropriate for your particular circumstances.
    Indeed, in some cases it's possible to cover all of the ``background''
    material either in the introduction or at appropriate places in
    the rest of the dissertation.

\end{guidance}

\section{Tracking Resources with Linearity}

\section{Problem in Detail}

\section{Proposed Solution}