### 2018-02-05 write-up/conclusion.tex

\chapter{Conclusion}

\begin{guidance}
    As you might imagine: summarizes the dissertation, and draws
    any conclusions. Depending on the length of your work, and
    how well you write, you may not need a summary here.

    You will generally want to draw some conclusions, and point
    to potential future work.
\end{guidance}