### 2018-02-05 write-up/declaration.tex

\newpage
{\normalfont\Huge\sffamily\bfseries Declaration}

\vspace{24pt}

I \authorname of \authorcollege, being a candidate for Computer Science Tripos,
Part III, hereby declare that this report and the work described in it are my
own work, unaided except as may be specified below, and that the report does
not contain material that has already been used to any substantial extent for a
comparable purpose.

\vspace{24pt}
Total word count: \wordcount

\vspace{60pt}
\textbf{Signed}:

\vspace{12pt}
\textbf{Date}:

\vfill

This dissertation is copyright \copyright 2018 \authorname. 
\\
All trademarks used in this dissertation are hereby acknowledged.

\newpage
\vspace*{\fill}