### 2018-02-05 write-up/introduction.tex

\chapter{Introduction}

% No resetting page numbers: ISO 7144
% \pagenumbering{arabic}
% \setcounter{page}{0}

\begin{guidance}
    This is the introduction where you should introduce your work.  In
    general the thing to aim for here is to describe a little bit of the
    context for your work --- why did you do it (motivation), what was the
    hoped-for outcome (aims) --- as well as trying to give a brief
    overview of what you actually did.

    It's often useful to bring forward some ``highlights'' into
    this chapter (e.g.\ some particularly compelling results, or
    a particularly interesting finding).

    It's also traditional to give an outline of the rest of the
    document, although without care this can appear formulaic
    and tedious. Your call.
\end{guidance}

\section{Overview of Problem}

\section{Contributions}