### 2018-02-05 write-up/related_work.tex

\chapter{Related Work}

\begin{guidance}

    This chapter covers relevant (and typically, recent) research
    which you build upon (or improve upon). There are two complementary
    goals for this chapter:
    \begin{enumerate}
      \item to show that you know and understand the state of the art; and
      \item to put your work in context
    \end{enumerate}

    Ideally you can tackle both together by providing a critique of
    related work, and describing what is insufficient (and how you do
    better!)

    The related work chapter should usually come either near the front or
    near the back of the dissertation. The advantage of the former is that
    you get to build the argument for why your work is important before
    presenting your solution(s) in later chapters; the advantage of the
    latter is that don't have to forward reference to your solution too
    much. The correct choice will depend on what you're writing up, and
    your own personal preference.

\end{guidance}