### 2018-04-19 write-up/background.tex
Start background and add matrices to semantics

\chapter{Background}

% \begin{guidance}
%     A more extensive coverage of what's required to understand your
%     work. In general you should assume the reader has a good undergraduate
%     degree in computer science, but is not necessarily an expert in
%     the particular area you've been working on. Hence this chapter
%     may need to summarize some ``text book'' material.
%
%     This is not something you'd normally require in an academic paper,
%     and it may not be appropriate for your particular circumstances.
%     Indeed, in some cases it's possible to cover all of the ``background''
%     material either in the introduction or at appropriate places in
%     the rest of the dissertation.
% \end{guidance}

\prechapter{%
    In this chapter, I will outline the concept of linear types and show how
    they can be used to solve the problems faced by programmers writing code
    using linear-algebra libraries. I will be emphasising the \emph{practical}
    and intuitive explanations of linear types to keep this thesis accessible
    to working programmers as well as academics not familiar with type-theory;
    giving only a terse overview of the history and theory behind linear types
    for the interested reader to pursue further.
}%

\section{Tracking Resources with Linearity}
Familiar examples of using a type-system to express program-invariants are
existential-types for abstraction and encapsulation, polymorphic types for
parametricity and composition (a.k.a generics). Less-known examples include
dependent-types (contracts or pre- and post-conditions). The advantages of
using a type-system to express program invariants are summarised by saying the
stronger the rules you follow, the better the guarantees you can get about your
program, \emph{before} you run it. At first, the rules seems restrictive, but
similar to how the rules of grammar, spelling and more generally writing help a
writer make it easier and clearer to communicate the ideas they wish to express,
so to do rules about typing make it easier to communicate the intent and
assumptions under which a program is written. An added, but often overlooked
benefit is automated-checking: a programmer can boldly refactor in certain
ways and the compiler will \emph{assist} in ensuring the relevant invariants
the type-system enforces are updated and kept consistent by pointing-out
where they are violated.

Linear types are a way to help a programmer track and manage resources. In
practical programming terms, they enforce the restriction that a value may be
used exactly once.\footnote{This definition may differ from more colloquial
uses in discussions surrounding \emph{substructural} type systems and/or Rust.
A brief explanation can be found in Section~\ref{sec:further}.} While this
restriction may seem limiting at first, precisely these constraints can be used
to express common invariants of the programs written by working programmers
every day. For example: a file or a socket, once opened \emph{must} closed; all
memory that is manually allocated must be freed.  C++'s destructors and Rust's
Drop-trait (and more generally, its borrow-checker) attempt to enforce these
constraints by basically doing the same thing: any resource that has not been
moved is deallocated at the end of the current lexical scope. Notably, these
languages also permit aliasing, alongside rules enforcing when it is acceptable
to do so.  On face value, the above one-line description of linear types
prevents aliasing or functions such as $\lambda x.\;x \times x$, such features
are still allowed (albeit in a more restricted fashion) in a \emph{usable}
linear type system designed for working with linear-algebra libraries.

\section{Problem in Detail}
Given this background, the most pertinent question at hand is: what problems do
linear-algebra library users (and writers) typically face? The answer to this
question depends on which of two buckets a programmer falls (or is forced by
domain) into. On one side, we have users of high-level linear-algebra libraries such
as Owl (for OCaml), Julia and Numpy (for Python); other the other, we have
users of more manual, lower-level libraries such as BLAS (Basic Linear Algebra
Subroutines) and Eigen for languages like C++ and FORTRAN.\footnote{I am not including
Rust in this comparison because its linear-algebra libraries are under active
development and not as well-known/used. Later on, given that it is a language
with in-built support of substructural features to track resources, Rust will
be compared and constrasted with this project to evaluate the classic
(E)DSL-versus-language-feature debate as it applies to the domain of
linear-algebra libraries.}
% https://github.com/AtheMathmo/rulinalg/blob/e0d8b76ddf8bbb5547cb9c98172e594605ab5fef/src/matrix/impl_ops.rs
% Which operand is overwritten determines scope. Would make sense to argue
% pros-cons of implicitness vs explicitness, emulating Rust's Affine with
% linear sugar.

\subsection{One Too Many Copies and a Thousand Bytes Behind}
% CODE EXAMPLES

\subsection{IHNIWTNM}
The title of this subsection\footnote{I Have No Idea What Those Names Mean.}
illustrates the problem with the C++/FORTRAN side: legibility (and ease of
reasoning) is sacrified at the alter of performance and efficiency.

% CODE EXAMPLES

\section{Proposed Solution}
A (E)DSL!

% CODE EXAMPLES

\section{Further Reading and Theory}\label{sec:further}
\begin{itemize}
    \item History - Girard and so on
    \item Congratulations, it's a co-monad
    \item Co-effects
    \item General direction of field - Haskell, Rust, RAML, Atkey/McBride \& Idris, Krishnaswami
\end{itemize}