### 2018-04-19 write-up/implementation.tex

\chapter{Implementation}

\begin{guidance}
This chapter may be called something else\ldots but in general
the idea is that you have one (or a few) ``meat'' chapters which
describe the work you did in technical detail.
\end{guidance}

\prechapter{%
    I implemented this project in OCaml, however I strongly believe the ideas
    described in this chapter can be applied easily to other languages and also
    are modular enough to extend the OCaml implementation to output to different
    back-end languages. I will show how a small core language with a few features
    can be the target of heavy-desugaring of typical linear-algebra programs. This
    core language can then be elaborated and checked for linearity before
    performing some simple and predictable optimisations and emitting (in this
    particular implementation) OCaml code that is not obviously safe, correct
    and performant.
}%

\section{Core Language}

\section{Elaboration and Inference}

\section{Compiling and Metaprogramming}