### 2018-05-08 write-up/conclusion.tex

\chapter{TO DO: Conclusion}

\begin{guidance}
    As you might imagine: summarizes the dissertation, and draws
    any conclusions. Depending on the length of your work, and
    how well you write, you may not need a summary here.

    You will generally want to draw some conclusions, and point
    to potential future work.
\end{guidance}

\textbf{What are the benefits of the type-based approach I have taken? Have I successfully 
argued whatever I wrote in the introduction?}

\section{Future Work}

\textbf{Graph stuff... check with supervisors.}

All of the matrix expression compilers mentioned in previous section construct
some sort of dataflow graph to represent the computation being executed. While
this seems intuitive, there is no formal argument for this  appr
Some directions in which a type-based approach to efficient matrix expression compilation could
be taken are:

\begin{itemize}

    \item \textbf{As a typed IR} for matrix expression compilers. This in turn
        could enable
        \begin{itemize}
            \item existing matrix expression compilers to be less opaque about
                what resources they are consuming.
            \item open up opportunities for non-local sharing of temporary values
                with some intra-procedural analysis.
            \item allow the user to choose: use a matrix expression compiler when
                desired and drop down to a usable typed-IR for finer control,
                whilst still retaining safety guarantees.
        \end{itemize}

    \item \textbf{Formal verification of matrix expression compilers} by
        precisely specifying source and target languages.

    \item \textbf{Multi-stage programming} to use information only available
        at runtime in many situations (such as sizes, matrix properties,
        sparsity, control flow) can be effectively incorporated into code
        generated.

    \item \textbf{Dependent types} to have control over how resources can be
        used and split. In addition to formal verification, dependent types
        could be combined with linearity to express finer-grain conventions
        surrounding blocking, slicing and writing to \emph{parts} of the matrix
        instead of the whole. This is already prevalent with `dsymm' like BLAS
        routines which only read and write to the lower or upper triangle of a
        matrix.  This idea is inspired by Conor McBride's talk on writing to
        terminals with ``Space Monads''~\cite{space_monads}.

    \item \textbf{Compiling to hardware} is also an option - once we know
        exactly when and where temporaries are required and what can be re-used
        when, we come one (small, but useful) step closer to realising matrix
        expressions directly on hardware.

\end{itemize}