### 2018-05-08 write-up/implementation.tex

\chapter{Implementation}

\begin{guidance}
This chapter may be called something else\ldots but in general
the idea is that you have one (or a few) ``meat'' chapters which
describe the work you did in technical detail.
\end{guidance}

\prechapter{%
    I implemented LT4LA in OCaml, however I strongly believe the ideas
    described in this chapter can be applied easily to other languages and also
    are modular enough to extend the OCaml implementation to output to different
    back-end languages. I will show how a small core language with a few features
    can be the target of heavy-desugaring of typical linear-algebra programs. This
    core language can then be elaborated and checked for linearity before
    performing some simple and predictable optimisations and emitting (in this
    particular implementation) OCaml code that is not obviously safe and correct
    (with respect to linearity).
}%

\section{Structure of LT4LA}

LT4LA follows the structure of a typical compiler for a (E)DSL\@. From the
start, I made a concerted effort to (1) write pure-functional code (typically
using a monadic-style) which helped immensely with modularity and debugging
when tests showed errors (2) produce readable, useful and precise
error-messages in the hope that someone who did not understand linear types
could still use the LT4LA (3) write tests and set-up
continuous-integration for all non-trivial functions so that I could spot and
correct errors that were not caught by OCaml's type-system whenever I
implemented new features or or refactored my code.

\begin{enumerate}

    \item \textbf{Parsing \& Desugaring}. A generated, LR(1) parser parses a text
        file into a syntax tree, which is then desugared into a smaller, more
        concise abstract syntax tree. The former aims to mimic OCaml syntax
        with a few extensions and keywords so that it is familiar and thus easy
        to pick-up for OCaml users. The latter allows for the type-checker to
        be simpler to implement and easier to specify. In general, this part
        will vary for different languages or can be dealt with differently
        using combinators (the EDSL approach) or a syntax-extension if the host
        language offers such support.

    \item \textbf{Type-checking}. The abstract syntax tree is explicitly-typed,
        with some inference to make it less verbose and more convenient to
        write typical programs.

    \item \textbf{Matrix Expressions}. During type-checking, if a matrix-expression
        is encountered, it is either successfully elaborated into an expression
        in the abstract syntax tree which is then consequently type-checked, or
        fails to find suitable routines to calculate the given expression.

    \item \textbf{Code Generation}. The abstract syntax tree is translated into
        standard OCaml and a few-particular `optimisations' are made to produce
        more readable code. This process is type-preserving: the linear type
        system is embedded into OCaml's type system, and so when the OCaml compiler
        compiles the generated code, it acts as a sanity check on the code produced.

    \item \textbf{Executable Artifacts}. A transpiler and a REPL are the main artifacts
        produced for this thesis.

    \item \textbf{Tests}. As mentioned before, almost all non-trivial functions
        have tests to check their behaviour. The output of the transpiler
        was also tested by having the build system generate code at OCaml code
        which could then be compiled and tested as usual.

\end{enumerate}

\section{Core Language}

A full description of the core language can be found in
Appendix~\ref{chap:ott_spec}. To make a linearly-typed language usable, we need
some way of using values zero or more than once, as we would an intuitionistic
value. For this, we have in the type-expressions the !-constructor and in the
term-expressions the \textbf{many}-constructor. For soundness, I implemented
\emph{value-restriction} for this type system, only \emph{values} (anything not
an array, matrix or an expression that can be reduced) that use \emph{no linear
values} can be marked with !. This means we can freely have integers, booleans
and (appropriate) functions usable as normal. Recursive functions can only use
linear values that are passed in as arguments to the function.

% TODO Once I've implemented the <- syntax, this should just include
% ../test/examples/weighted_avg_infer.lt directly
\begin{figure}[tb]
    \begin{minted}[linenos]{ocaml}
let rec f (!i : !int)
        (!n : !int)
        (!x0 : !elt)
        (write : z arr)
        ('x)
        (weights : 'x arr)
        : z arr * 'x arr =
  if n = i then (write, weights) else
  let !w0 <- weights[0] in
  let !w1 <- weights[1] in
  let !w2 <- weights[2] in
  let !x1 <- write[i] in
  let !x2 <- write[i + 1] in
  let !x : !elt = w0 *. x0 +. w1 *. x1 +. w2 *. x2 in
  f (i + 1) n x1 (write[i] := x) _ weights
    \end{minted}

    \caption{A recursive function to perform a \emph{in-place} (simplified)
        1D-convolution (sliding weighted-average) using a read-only array of
        weights and a writeable array, originally containing the input
        array.}\label{fig:oned_conv}

\end{figure}

So, how does a typical programmer write code that uses a matrix or an array
more than once? Figure~\ref{fig:oned_conv} shows an example program that
executes a (simplified) 1D-convolution (sliding weighted-average) using a
read-only array of three weights. On line 1, \ltfla{let rec} introduces a
recursive function `f', followed by explicitly annotated formal parameters.
Intuitively, \ltfla{!id} allows \ltfla{id} to be used intuitionistically. The
types \ltfla{!int} and \ltfla{!elt} correspond to integers and element types
(for example, 64-bit floating-point).

Array and matrix types are parameterised by \emph{fractional-capabilities}.  A
fraction of 1 ($2^0$) represents complete ownership of a value; in
particular, this allows a programmer to write or free it. Creating an array
gives you ownership of it; the function \ltfla{array : !int --o z arr}
(where \ltfla{z} represents `0'). Once you have ownership of an
array/matrix, you can free it: \ltfla{free : z arr --o unit}.
Importantly, because a linear-value may only be used once, the array just freed
is \emph{out of scope} for following expressions, preventing use-after-free.
Ownership also enables you to write to the array: \ltfla{set : z arr
--o !int --o !elt --o z arr} (the syntax \ltfla{w[i] := j} is just
sugar for \ltfla{set w i j}). Here, linearity prevents accessing values
which represented the array \emph{before} the mutation.

Any fraction less than 1 (for simplicity, limited to $2^{-k}$ in this system,
for a positive integer $k$) represents read-only access. So, the \ltfla{'x}
represents a natural number (either a zero \ltfla{z}, variable \ltfla{'x} or a
successor (+1) of a natural number). Hence, you can read from (index) any array
\ltfla{get : 'x .  'x arr --o !int --o !elt * 'x arr} (the syntax \ltfla{let !v
<- w[i]} is just sugar for \ltfla{let (!v, w) = get _ w i}). Transparently
rebinding \ltfla{w} with the returned value means a program can appear to use
\ltfla{w} multiple times; the underscore is how a programmer tells the compiler
to automatically \emph{infer} the correct fractional-capability based on the
other arguments passed to the function. Fractions also permit aliasing:
\ltfla{share : 'x . 'x arr --o 'x s arr * 'x s arr}. However, given the types
of the primitives provided, these aliases can only be read from and not written
to. If you want to write to this array, you must \ltfla{unshare: 'x . 'x s arr
--o 'x s arr --o 'x arr} it until you are left with a value of type \ltfla{z
arr}, guaranteeing no other aliases exists.

Given this set-up, we now \emph{statically} have \emph{perfect} information
about aliasing and ownership of values in the program. We can only write to an
array when we own it; ownership guarantees no other aliases exist in scope at
the point of usage. In the next section, I show how this perfect information
can be used to write more natural-looking code using value-semantic expressions
which behave in precisely the way we intend it to. Now the programmer need not
resort to manually tracking and inserting \texttt{noalias} annotations; instead
they can let the loyal and tireless compiler do the heavy lifting.

\section{Matrix Expressions}

We have now arrived at the titular \emph{applications} of linear types. I will
show how, in addition to preventing the errors explained in the previous
section, we can take linear types one step further and apply it to the domain
of efficient compilation of matrix-expressions.

\begin{figure}[tb]
    \inputminted[linenos]{fortran}{kalman.f90}
    \caption{Kalman filter in Fortran 90 (Credit: Matthew
        Rocklin~\cite{rocklin_gist}).}\label{fig:fortran_kalman}
\end{figure}

In Figure~\ref{fig:fortran_kalman}, we see the difficulty of efficiently
implementing a \emph{Kalman filter}, a powerful and set of equations applicable
to a wide variety of problems. From the comments, we see that every variable is
annotated with the step/matrix expression that it will hold at some point
during the computation (an equivalent alternative, say in C++, could be to have
a meaningful name for each step/matrix expression and manually annotate/keep
track of which names alias the same location).

In contrast, Figure~\ref{fig:ltfla_kalman}, offers the advantages of
\begin{itemize}
    \item aliasing: labelling each step with a different, more meaningful variable name,
    \item easily spotting which resources are being passed in and which are
        allocated for the function (new/copy),
    \item unambiguously seeing \emph{when} and what values are freed;
\end{itemize}
and have the compiler automatically ensure the safety of each of the above by respectively
\begin{itemize}
    \item making it impossible to refer to steps/values which are no longer usable,
    \item ensuring all values are declared and \emph{initialised} correctly before they are used,
    \item checking no values are used after they are freed \emph{and} ensuring no values are leaked.
\end{itemize}

Indeed, an inexperienced programmer could take the  na\"ive approach of just
copying sub-expressions by default and then letting the compiler tell it which
copies are never used and removing them systematically until it type checks.
While it is not quite a black-box, push-button compilation of an expression, I
would argue that, it is just as easy (if not easier) to become familiar with as
Rust and its borrow-checker.

\begin{figure}[tb]

    \begin{minted}[linenos]{ocaml}
let !kalman
        ('s) (sigma : 's mat) (* n, n *)
        ('h) (h : 'h mat)     (* k, n *)
        ('m) (mu : 'm arr)    (* n    *)
        (r_1 : z mat)         (* k, k *)
        (data_1 : z arr)      (* k    *)
        : ('s mat * ('h mat * 'm arr)) * (z arr * z mat) =
    let ((!k, !n), h) = dimM _ h in
    let zeros = zeroM n n in
    let sigma_h = new [| sym (up, sigma) * h |] in
    let r_2 = [| h * sigma_h + r_1 |] in
    let mu_1 <- copy mu in
    let sigma_1 <- copyM sigma in
    let data_2 = [| h * mu + -1. * data_1 |] in
    let h_1 <- copyM h in
    let h_sol = [| sym (up, r) ^ -1 * h_1 ] in
    let data_sol = [| sym (up, r) ^ -1 * data_2 |] in
    let h_h1 = new [| h * h_sol |] in
    let h_data = new [| h * data |] in
    let h_h1_sigma = new [| sym (up, h_h1) * sigma |] in
    let mu_2 = [| sym (up, sigma) * h_data * mu_1 |] in
    (* non-linear use of sigma_c..? *)
    let sigma_c = [|  -1. * sym (up, sigma_c) * h_h1_sigma + sigma_c |] in
    (* frees... *)
    (mu_c , sigma_c)
    \end{minted}

    \caption{Kalman filter in LT4LA.}\label{fig:ltfla_kalman}

\end{figure}

\subsection{TO DO: Elaboration}

So-called `matrix \emph{expressions}' are still `side-effecting'/consuming
linear values, and do not produce \emph{values} but a \emph{sequence of
(re-)bindings} which in turn dictate what values are still available/in-scope
after the computation.  As such, compilation cannot be done purely
compositionally via structural-induction on the expression language; it is most
concisely expressed via CPS, where the result of elaborating a matrix
expression is a function that takes as its argument the rest of the computation
expecting to use the result it has just made available under that requested
name.

\subsection{TO DO: Inference}

\section{TO DO: Code Generation}

High-level overview: primitives, translations, optimisations, difficulty, build system.

\subsection{TO DO: Compiling Constructs to OCaml}

\subsection{TO DO: Metaprogramming}

\section{TO DO: Theory}

Fractional permissions, not proven soundness etc\. but compilation preserves
types as checked by OCaml.