### 2018-05-22 write-up/implementation.tex

\chapter{Implementation}\label{chap:impl}

% \begin{guidance}
% This chapter may be called something else\ldots but in general
% the idea is that you have one (or a few) ``meat'' chapters which
% describe the work you did in technical detail.
% \end{guidance}

\prechapter{%
    I implemented LT4LA in OCaml, however I strongly believe the ideas
    described in this chapter can be applied easily to other languages and also
    are modular enough to extend the OCaml implementation to output to different
    back-end languages. I will show how a small core language with a few features
    can be the target of heavy-desugaring of typical linear-algebra programs. This
    core language can then be elaborated and checked for linearity before
    performing some simple and predictable optimisations and emitting (in this
    particular implementation) OCaml code that is not obviously safe and correct
    (with respect to linearity).
}%

\section{Structure of LT4LA}

LT4LA follows the structure of a typical compiler for a (E)DSL\@. From the
start, I made a concerted effort to (1) write pure-functional code (typically
using a monadic-style) which helped immensely with modularity and debugging
when tests showed errors (2) produce readable, useful and precise
error-messages in the hope that someone who did not understand linear types
could still use the LT4LA (3) write tests and set-up
continuous-integration for all non-trivial functions so that I could spot and
correct errors that were not caught by OCaml's type-system whenever I
implemented new features or or refactored my code.

\begin{enumerate}

    \item \textbf{Parsing \& Desugaring}. A generated, LR(1) parser parses a text
        file into a syntax tree, which is then desugared into a smaller, more
        concise abstract syntax tree. The former aims to mimic OCaml syntax
        with a few extensions and keywords so that it is familiar and thus easy
        to pick-up for OCaml users. The latter allows for the type-checker to
        be simpler to implement and easier to specify. In general, this part
        will vary for different languages or can be dealt with differently
        using combinators (the EDSL approach) or a syntax-extension if the host
        language offers such support.

    \item \textbf{Type-checking}. The abstract syntax tree is explicitly-typed,
        with some inference to make it less verbose and more convenient to
        write typical programs.

    \item \textbf{Matrix Expressions}. During type-checking, if a matrix-expression
        is encountered, it is either successfully elaborated into an expression
        in the abstract syntax tree which is then consequently type-checked, or
        fails to find suitable routines to calculate the given expression.

    \item \textbf{Code Generation}. The abstract syntax tree is translated into
        standard OCaml and a few-particular `optimisations' are made to produce
        more readable code. This process is type-preserving: the linear type
        system is embedded into OCaml's type system, and so when the OCaml compiler
        compiles the generated code, it acts as a sanity check on the code produced.

    \item \textbf{Executable Artifacts}. A transpiler and a REPL are the main
        artifacts produced for this thesis. For evaluation, I implemented
        Kalman filters in Owl, LT4LA and CBLAS/LAPACKE and a benchmarking
        program to measure execution times.

    \item \textbf{Tests}. As mentioned before, almost all non-trivial functions
        have tests to check their behaviour. The output of the transpiler was
        also tested by having the build system generate OCaml code at compile
        time, which in turn could then be compiled and tested like handwritten
        OCaml code.

\end{enumerate}

\section{Core Language}

A full description of the core language can be found in
Appendix~\ref{chap:ott_spec}. Its main features are intuitionistic values,
value-restriction, fractional-capabilities (inferrable at call sites),
if-expressions and recursion.

\subsection{Intuitionistic Values}

To make a linearly-typed language usable, we need some way of using values zero
or more than once, as we would an intuitionistic value. For this, we have in
the type-expressions the !-constructor and in the term-expressions the
\ltfla{Many}-constructor and the \ltfla{let Many <id> = .. in .. } eliminator.
The idea behind the !-type is that the value uses no `resources'
(linearly-typed expressions). To start off with, it is enough to say that
anything which can be passed around by copying, will have a !-type. This
includes integers, elements and booleans. So \ltfla{3 : !int} and \ltfla{3. +.
4. : !elt}. However, all bindings are still linear by default, so to emulate
intuitionism, I desugar \ltfla{let !x = <exp> in <body>} to  \ltfla{let Many x
= <exp> in let Many x = Many (Many x) in <body>} (similarly for function
argument bindings). The reader can check (using the rules in
Appendix~\ref{chap:ott_spec}) that this has the effect of moving \ltfla{x : !t}
from the linear to the intuitionistic environments, only if \ltfla{<exp> : !t}.

However, just that desugaring alone is not enough to prevent a user from taking
an array or matrix and moving it into the intuitionistic environments. Why?
There are certain situations in which we \emph{should not} use the \ltfla{Many}
constructor.  Consider the following code: \ltfla{let Many x = Many (array 5)
in <body>}; the expression \ltfla{array 5} uses no linearly-typed variables
from the linear environment. Although we could just reject types of the form
\ltfla{!(_ arr)} to fix this simple example, what about pairs \ltfla{let Many
xy = Many (3, array 5) in <body>}?  Ad-hoc pattern matching on the type cannot
account for all possible situations. With the last case, we can use \ltfla{xy}
as many times as we would like, destruct the pair to get the second component
and thus create \emph{distinct} read-write aliases to the same array.  Alas,
now arrays can be used intuitionistically and all the benefits of linearity are
lost.  Or are they?

\subsection{Value-Restriction}

Not quite, but to understand how we can fix this problem, we need to question
an assumption left implicit up until this point: what does \ltfla{Many} even
mean? What does it do at runtime? One option is to go down the C++ route and
make \ltfla{Many} act like a \ltfla{shared_ptr} and act as a runtime
reference-count for arrays. I chose to not go for this option because it went
against the \emph{explicitness} and \emph{predictability} that C and Fortran
have. It would make analysing when and what is allocated and freed more like
the higher-level languages I was trying to move away from.

My aim is to show linear types can be simple to understand and apply to linear
algebra, enough so that it can be grafted (in a limited way) on to existing
languages as a \emph{library}. In that spirit, the simplest thing that
\ltfla{Many} can mean at runtime is \emph{nothing}. This language construct is
translated into a standard OCaml language constructor of the form \ltfla{type
'a bang = Many of 'a [@@unboxed]}. The \emph{unboxed} annotation means
that the type and its constructor only exist for the purpose of type-checking
in OCaml; \emph{the runtime representation of values of type \ltfla{'a} and
\ltfla{'a bang} is exactly the same}.

With this understanding, our problem is that arrays and matrices are unlike
other values such as integers and elements because (under the OCaml hood)
calling a function with an array argument copies a \emph{pointer} to the array
rather than the array itself, instead of the \emph{value} itself. So, we can
start making a distinction, \emph{defining} elements, integers, booleans,
intuitionistic variables, units and lambda-expressions that capture no linear
variables as \emph{values} (since they cannot break referential-transparency)
and anything else (arrays, matrices, expressions which can be reduced, such as
function application and if-expressions) as not being `values'. If this sounds
familiar, it is because this is the same \emph{value-restriction} `trick' from
the world of polymorphic types applied to linearity instead. We then have the
rule that we can only use \ltfla{Many} on expressions that are defined to be
\emph{values} and \emph{use no linear variables}.

\subsection{Fractional-Capabilities and Inference}

Having started off with linearity and understanding how it helps a programmer
keep track of of memory allocation and deallocation, and then understood and
correctly implemented intuitionism for the only values we wish to be
intuitionistic, and, we now tackle the problem of how to implement safe
aliasing and read/write permissions.

Array and matrix types are parameterised by \emph{fractional-capabilities}.  A
fraction of 1 ($2^0$) represents complete ownership of a value; in particular,
this allows a programmer to write or free it. Creating an array gives you
ownership of it; the function \ltfla{array : !int --o z arr} (where \ltfla{z}
represents `0'). Once you have ownership of an array, you can free it:
\ltfla{free : z arr --o unit}.  Importantly, because a linear-value may only be
used once, the array just freed is \emph{out of scope} for following
expressions, preventing use-after-free.  Ownership also enables you to write to
the array: \ltfla{set : z arr --o !int --o !elt --o z arr} (the syntax
\ltfla{w[i] := j} is just sugar for \ltfla{set w i j}). Here, linearity
prevents accessing aliases which represented the array \emph{before} the
mutation.

Any fraction less than 1 (for simplicity, limited to $2^{-k}$ in this system,
for a positive integer $k$) represents read-only access. So, the \ltfla{'x}
represents a natural number (either a zero \ltfla{z}, variable \ltfla{'x} or a
successor (+1) of a natural number). Hence, you can read from (index) any array
\ltfla{get : 'x .  'x arr --o !int --o 'x arr * !elt} (the syntax \ltfla{let !v
<- w[i]} is just sugar for \ltfla{let (w, !v) = get _ w i}). In general, a
left-arrow \ltfla{<-} signifies transparent rebinding with returned values: it
means a program can \emph{appear} to use a variable multiple times, important
for keeping LT4LA usable and readable. The underscore is how a programmer tells
the compiler to automatically \emph{infer} the correct fractional-capability
based on the other arguments passed to the function. In conjunction with the
requirement that functions declarations need type-annotations for their
arguments, this allows a fractional-capability to be correctly inferred in any
program.

Fractions exist to provide both safe aliasing and read/write permissions, via
the primitives \ltfla{share : 'x . 'x arr --o 'x s arr * 'x s arr} and
\ltfla{unshare: 'x .  'x s arr --o 'x s arr --o 'x arr}.  For the former, two
arrays returned (which happen to just be the given array) can now only be read
from and not written to. If you want to write to this array, you must use the
latter to combine other read-only aliases until you are left with a value of
type \ltfla{z arr}, guaranteeing no other aliases exists.

Given this set-up, we now \emph{statically} have \emph{perfect} information
about aliasing and ownership of values in the program. We can only write to or
free an array only when we own it; ownership guarantees no other aliases exist
in scope at the point of usage. In Figure~\ref{fig:ltfla_kalman}, I show how
this perfect information can be used to write more natural-looking code using
value-semantic expressions which behave in precisely the way we intend it to.
Now the programmer need not resort to manually figuring out and inserting
\texttt{noalias} annotations and worrying about what variables can and cannot
be written to or freed; instead they can let the loyal and tireless compiler do
the heavy lifting.

\subsection{If-Expressions}

Because we do not know which way a condition will evaluate at run time, we must
guarantee the both branches use the same set of linear variables. Writing the
type-checker in a pure-functional monadic style paid off here because I could
now sandwich monadic values with state-adjustments either side of it. Given two
monadic values that represented type-checking two branches of an if-expression,
I could use the code in Figure~\ref{fig:same_resources} to easily save, reset
and compare the state either side of running those monadic values.

\begin{figure}[tp]
    \begin{minted}[fontsize=\small]{ocaml}
let same_resources (wf_a, loc_a) (wf_b, loc_b) =
  let open Let_syntax in
  (* Save state *)
  let%bind {used_vars=prev; env=old_env; _} as state = get in
  (* Reset, run a, save state *)
  let%bind () = put { state with used_vars = empty_used } in
  let%bind res_a = wf_a in
  let%bind {used_vars=used_a; _} as state = get in
  (* Reset, run b, save state *)
  let%bind () = put { state with used_vars = empty_used; env = old_env } in
  let%bind res_b = wf_b in
  let%bind {used_vars=used_b; _} as state = get in
  (* Check if same resources *)
  let keys_a, keys_b = (* convert to (used_a, used_b) to sets *) in
  if Set.equal keys_a keys_b then
    (* merge used_vars and used_b environments *)
  else
    (* report differences *)
    \end{minted}

    \caption{Implementation of same\_resources helper method for type-checking
        if-expressions. Note the monadic style helped compose computations that
        affected the type-checker's state in a simple manner.}\label{fig:same_resources}

\end{figure}

\subsection{Functions and Recursion}

A non-recursive function may be used more than once if it does not refer to any
linear variables from the surrounding scope. So, we can desugar something like
\ltfla{let x = 3 in let !f (y : !int) = x + y in <body>} to \ltfla{let x = 3 in
let Many f =  Many (fun y : !int -> x + y) in  <body>}. Recursion is
slightly more complicated: we can desugar the factorial function
(Figure~\ref{fig:ltfla_factorial}) to \ltfla{let Many f = fix (f, x : !int,
if (*..*) : !int) in <body> }. However, \ltfla{fix}, like \ltfla{Many},
must also not use any linear variables from its surrounding scope.

\section{Matrix Expressions}

We have now arrived at an extended application of linear types. I will show
how, we can apply the ability to automatically check aliasing, read/write
permissions, memory allocation, re-use and deallocation, to the domain of
matrix-expression compilation.

\begin{sidewaysfigure}
    \inputminted[linenos, fontsize=\footnotesize]{fortran}{kalman.f90}
    \caption{Kalman filter in Fortran 90.}\label{fig:fortran_kalman}
\end{sidewaysfigure}

In Figure~\ref{fig:fortran_kalman}, we see the difficulty of efficiently
implementing a \emph{Kalman filter}, a powerful set of equations applicable
to a wide variety of problems. From the comments, we see that every variable is
annotated with the step/matrix expression that it will hold at some point
during the computation (an equivalent alternative, say in C++, could be to have
a meaningful name for each step/matrix expression and manually annotate/keep
track of which names alias the same location).

In contrast, Figure~\ref{fig:ltfla_kalman}, offers the advantages of
\begin{itemize}
    \item aliasing: labelling each step with a different, more meaningful variable name,
    \item easily spotting which resources are being passed in and which are
        allocated for the function (new/copy),
    \item unambiguously seeing \emph{when} and what values are freed;
\end{itemize}
and have the compiler automatically ensure the safety of each of the above by respectively
\begin{itemize}
    \item making it impossible to refer to steps/values which are no longer usable,
    \item ensuring all values are declared and \emph{initialised} correctly before they are used,
    \item checking values are neither used after they are freed \emph{nor} leaked.
\end{itemize}

Indeed, an inexperienced programmer could take the  na\"ive approach of just
copying sub-expressions by default and then letting the compiler tell it which
copies are never used and removing them systematically until it type-checks.
While it is not quite a black-box, push-button compilation of an expression, I
would argue that, it is just as easy (if not easier) to become familiar with as
Rust and its borrow-checker.

\begin{figure}[tp]
    \inputminted[linenos, fontsize=\small]{ocaml}{../test/examples/kalman.lt}
    \caption{Kalman filter in LT4LA.}\label{fig:ltfla_kalman}
\end{figure}

\subsection{Elaboration}

All of the syntax in Figure~\ref{fig:ltfla_kalman} can be unambiguously
desugared \emph{before} type-checking, through fairly simple pattern matching.
Matrix expression compilation is well-trodden territory in
academia~\cite{rocklin_thesis, fabregat_thesis, gunnels_flame, linnea, taco}
but this is, to my knowledge, the first type-based approach to it.

An overview of the translations are in Figure~\ref{fig:mat_patterns}; details
about choosing between `symm' or `gemm' are omitted for brevity. Before
settling on this approach, I tried implementing a more general type-directed,
nested matrix expression compiler; I will now highlight some of the
difficulties inherent in the problem.

\begin{figure}[tp]
    \begin{minted}[fontsize=\small]{ocaml}
let x <- [| y |] in ..              ==> let (y,x) = copyM_to _ y x in ..
let x <- new [| y |] in ..          ==> let (y,x) = copyM _ y in ..
let x <- [| a*b + c |] in ..        ==> let x <- [| 1.*a*b + 1.*c |] in ..
let x <- [| i.*a*b + j.*c |] in ..  ==> let ((a,b), c) = (* BLAS *) in ..
let x <- new (m, n) [| a*b |] in .. ==> let c = matrix m n in
                                        let x <- [| 1.*a*b + 0.*c |] in ..
    \end{minted}
    \caption{Syntactic translations of matrix expressions to linearly-typed
        matrix functions. Further annotations on the matrix variables
        (\ltfla{sym} or \ltfla{^T}) determine which BLAS routine and what
        parameters to call it with.}\label{fig:mat_patterns}

\end{figure}

One of the first hurdles I encountered was compositionality: to compile $AB +
C$ it is typically better to use a BLAS routine directly rather than but first
compiling $A$, then $B$, then adding a call to multiply them, then $C$ and
finishing with a call to add the results.

Another compositionality problem is that a call to a linear function does not
just return a result, but a sequence of re-bindings that dictate which
variables are still in scope/usable.  As such, to compile an expression, you
need to provide a CPS-style function of type $var \rightarrow exp$ representing
how you would use a variable representing the result in the rest of the
expression.

However, the type of that variable also determines how you can use it: can you
write to it or must you copy it? This information depends on how the expression
representing the variable was elaborated and adds to the complexity of the
pattern-matching.

Copying leads us into dealing with temporaries: do you first allocate all
temporaries in new matrices and then analyse lifetimes and dimensions to figure
out which slots of memory can be re-used? Or do you try and infer the available
resources from the environment as you go?  Can, and should, you type arrays and
matrices (or $n$-dimensional tensors) as the same?

Adding in more and more considerations, the problem starts to resemble register
allocation: there are registers of different types and sizes, many instructions
to choose from, a cost model to take into account all whilst trying to balance
the number of registers in use and the number of instructions emitted.

\section{Code Generation}

Code generation is a thankfully straightforward mapping from core LT4LA
constructs to OCaml constructs, with the appropriate sprinkling of \ltfla{Many}
constructors for integer, element and boolean literals.

To make the code produced readable, I added a few `optimisations', to the
compiler, to `re-sugar' some of the constructs translated where appropriate.
Almost all of them involved the erasing the !-eliminator which is not needed in
regular, intuitionistic OCaml. So, so we can simply replace any expression of
the form:

\begin{itemize}

    \item \ltfla{let xy, z = <exp> in let x,y = xy in <body>} with
        \ltfla{let (x,y), z = <exp in <body>},

    \item \ltfla{let Many x = x in let Many x = Many (Many x) in <body>} with
        \ltfla{<body>},

    \item \ltfla{let Many x = <exp> in let Many x = Many (Many x) in <body>} with
        \ltfla{let x = <exp> in <body>},

    \item \ltfla{let Many x = Many <exp> in <body>} with \ltfla{let x = <exp> in <body},

    \item \ltfla{let Many f = fix (f, x, .., <exp>, ..) in <body>} with
        \ltfla{let rec f x = <exp> in <body>}.

\end{itemize}

The end result is visible in Figures~\ref{fig:ocaml_sumarray}
and~\ref{fig:ocaml_kalman}: OCaml code that is not obviously safe and correct
with respect to linearity, as promised at the start of the chapter.

\begin{figure}[tp]
    \begin{minted}[linenos, fontsize=\small]{ocaml}
let rec f i n x0 row =
  if Prim.extract @@ Prim.eqI i n then (row, x0)
  else
    let row, x1 = Prim.get row i in
    f (Prim.addI i (Many 1)) n (Prim.addE x0 x1) row
in
f
    \end{minted}
    \caption{Recursive OCaml function for a summing an array, generated from
        Figure~\ref{fig:ltfla_sumarray}, passed through \texttt{ocamlformat} for
        presentation.}\label{fig:ocaml_sumarray}

\end{figure}

Even though Figure~\ref{fig:ocaml_kalman} is in OCaml, I see little reason that
it cannot be C or Fortran. Translating LT4LA's core language constructs to
those languages would not be as easy as translating to OCaml. However, it is
easy to see that both Figures~\ref{fig:fortran_kalman}
and~\ref{fig:ocaml_kalman} are realisations of a concise, linearly-typed Kalman
filter~\ref{fig:ltfla_kalman} \emph{specification} that describes the whole
behaviour and intent (with respect to read/writes permissions, memory
management and aliasing) of the program and the BLAS primitives it uses more
accurately than a Fortran, C or OCaml implementation could.

\begin{figure}[p]
    \begin{minted}[linenos, fontsize=\footnotesize]{ocaml}
let kalman sigma h mu r_1 data_1 =
  let h, _p_k_n_p_ = Prim.size_mat h in
  let k, n = _p_k_n_p_ in
  let sigma_h = Prim.matrix k n in
  let (sigma, h), sigma_h =
    Prim.symm (Many true) (Many 1.) sigma h (Many 0.) sigma_h
  in
  let (sigma_h, h), r_2 =
    Prim.gemm (Many 1.) (sigma_h, Many false) (h, Many true) (Many 1.) r_1
  in
  let (h, mu), data_2 =
    Prim.gemm (Many 1.) (h, Many false) (mu, Many false) (Many (-1.)) data_1
  in
  let h, new_h = Prim.copy_mat_to h sigma_h in
  let r_2, new_r = Prim.copy_mat r_2 in
  let chol_r, sol_h = Prim.posv new_r new_h in
  let chol_r, sol_data = Prim.potrs chol_r data_2 in
  let () = Prim.free_mat chol_r in
  let h_sol_h = Prim.matrix n n in
  let (h, sol_h), h_sol_h =
    Prim.gemm (Many 1.) (h, Many true) (sol_h, Many false) (Many 0.) h_sol_h
  in
  let () = Prim.free_mat sol_h in
  let h_sol_data = Prim.matrix n (Many 1) in
  let (h, sol_data), h_sol_data =
    Prim.gemm (Many 1.) (h, Many true) (sol_data, Many false) (Many 0.) h_sol_data
  in
  let mu, mu_copy = Prim.copy_mat mu in
  let (sigma, h_sol_data), new_mu =
    Prim.symm (Many false) (Many 1.) sigma h_sol_data (Many 1.) mu_copy
  in
  let () = Prim.free_mat h_sol_data in
  let h_sol_h_sigma = Prim.matrix n n in
  let (sigma, h_sol_h), h_sol_h_sigma =
    Prim.symm (Many true) (Many 1.) sigma h_sol_h (Many 0.) h_sol_h_sigma
  in
  let sigma, sigma_copy = Prim.copy_mat_to sigma h_sol_h in
  let (sigma, h_sol_h_sigma), new_sigma =
    Prim.symm (Many false) (Many (-1.)) sigma h_sol_h_sigma (Many 1.) sigma_copy
  in
  let () = Prim.free_mat h_sol_h_sigma in
  ((sigma, (h, (mu, (r_2, sol_data)))), (new_mu, new_sigma)) )
in
kalman
    \end{minted}
    \caption{OCaml code for a Kalman filter, generated from
        Figure~\ref{fig:ltfla_kalman}, passed through \texttt{ocamlformat} for
        presentation.}\label{fig:ocaml_kalman}

\end{figure}

\subsection{Build System}

So once you have written your memory-optimised program with all the features
and support provided by LT4LA and made it produce well-typed, compilable OCaml
code, the question then becomes, how to use this code. This process will vary
across ecosystems, but within OCaml, the new build-system on the block
\emph{Dune} has support for generating, compiling and linking OCaml modules at
\emph{compile time}. Indeed, this is how I have written tests and benchmarks
for programs produced by LT4LA from within OCaml. In particular, I can use my
linearly-typed Kalman filter implementation just like and with any other OCaml
function.

Advantages of generating code at compile time include avoiding runtime overhead
as well as catching any interface changes between the generated code and the
code that uses it via OCaml's type system (another benefit of embedding LT4LA's
type system inside of OCaml).

I suspect that this approach will be very valuable to not just users of
libraries such as Numpy or Owl, but also their implementors: library functions
which presented a safe, value-semantic interface but use unsafe, mutating
operations on the inside could now be expressed using LT4LA and have an extra
safety-net and automatic checking for their implementations.

\section{Summary}

I have explained how a few core features -- linearity, the \ltfla{Many}
constructor, value-restriction, fractional-capabilities with inference,
if-expressions and recursive functions -- are enough to \emph{statically
capture and automatically check} aliasing, read/write permissions, memory
allocation, re-use and deallocation of real world linear algebra library
routines. I have also demonstrated that simple pattern-matching and desugaring
provides the potential for a new, \emph{type-directed} approach to matrix
expression compilation. Lastly I have shown that it is possible to use these
features with \emph{existing} languages and frameworks.\footnote{As mentioned
in the previous chapter, if the host language supports
\emph{syntax-extensions}, like PPX for OCaml, it would be possible to use LT4LA
features \emph{from within} OCaml instead of just alongside it.}