### 2018-05-22 write-up/introduction.tex

\chapter{TO DO: Introduction}

% No resetting page numbers: ISO 7144
% \pagenumbering{arabic}
% \setcounter{page}{0}

% \begin{guidance}
%     This is the introduction where you should introduce your work.  In
%     general the thing to aim for here is to describe a little bit of the
%     context for your work --- why did you do it (motivation), what was the
%     hoped-for outcome (aims) --- as well as trying to give a brief
%     overview of what you actually did.
% 
%     It's often useful to bring forward some ``highlights'' into
%     this chapter (e.g.\ some particularly compelling results, or
%     a particularly interesting finding).
% 
%     It's also traditional to give an outline of the rest of the
%     document, although without care this can appear formulaic
%     and tedious. Your call.
% \end{guidance}

\prechapter{%
}%

\section{Overview of Problem}

\section{Contributions}