### 2018-05-24 write-up/eval_data.tex

\chapter{Evaluation Raw Data}\label{chap:eval_data}

Below is formatted output from a trace I obtained by recompiling Owl with print
statements inserted on the relevant primitives. I made two modifications: I
shortened `\_matrix\_transpose' to `\_mtrsp' for formatting and I split `posv'
into `potrf/potrs' for a fairer comparison.

\inputminted[linenos, fontsize=\small]{text}{trace.txt}

Below is the raw output from the benchmarking script I wrote.
\inputminted[linenos, fontsize=\small]{text}{timings.txt}