### 2018-05-24 write-up/abstract.tex
Finish Evaluation

\newpage
{\normalfont\Huge\sffamily\bfseries TO DO: Abstract}
\vspace{24pt}

In this thesis, I argue that linear types are an appropriate, \emph{type-based
formalism} for expressing aliasing, read/write permissions, memory allocation,
re-use and deallocation, first, in the context of the APIs of linear algebra
libraries and then in the context of matrix-expression compilation.

I show that framing the problem using linear types can help \emph{reduce bugs}
by making precise and explicit the informal, ad-hoc practices typically
employed by human experts and linear algebra \emph{compilers} and automate
checking them.

As evidence for this argument, I show non-trivial, yet readable, linear algebra
programs, that are safe (with respect to aliasing, read/write permissions,
memory re-use and deallocation) and explicit (with respect to memory
allocation) which (1) are more memory-efficient than equivalent programs
written using high-level linear algebra libraries and (2) perform just
as predictably as equivalent programs written using low-level linear-algebra
libraries. I also argue \emph{the experience} of writing such programs with
linear types is qualitatively better in key-respects. In addition to all of
this, I show that it is possible to provide such features \emph{as a library}
on top of existing programming languages and linear algebra libraries.

\textbf{Summarise evaluation.}

\textbf{Summary future work.}

\newpage
\vspace*{\fill}