### 2018-05-27 write-up/implementation.tex

\chapter{Implementation}\label{chap:impl}

% \begin{guidance}
% This chapter may be called something else\ldots but in general
% the idea is that you have one (or a few) ``meat'' chapters which
% describe the work you did in technical detail.
% \end{guidance}

\prechapter{%
    I will describe the structure of LT4LA: a DSL (domain specific language)
    \textbf{library} written in OCaml that transpiles to OCaml. I will explain
    the features of its core language and show how readable (equational and
    algebraic expressions) and easy to reason about linear-algebra programs
    can be elaborated into memory-efficient ones in the core language that are
    then checked for safety (with respect to aliasing, read/write permissions,
    memory allocation, re-use and de-allocation). Finally, I will explain how
    such programs can be transpiled to OCaml code that is not obviously safe.
    Although I refer to OCaml-specific features of the type-checker and
    transpiler, I believe the ideas described in this chapter can easily be
    implemented in other languages and are also general enough to transpile to
    different back-end languages, such as C or Fortran.
}%

\section{Structure of LT4LA}\label{sec:structure_lt4la}

LT4LA follows the structure of a typical compiler for a (E)DSL\@. From the
start, I made a concerted effort to (1) write pure-functional code (typically
using a monadic-style) which helped immensely with modularity and debugging
when tests showed errors (2) produce readable, useful and precise
error-messages in the hope that someone who did not understand linear types
could still use the LT4LA (3) write tests and set-up
continuous-integration for all non-trivial functions so that I could spot and
correct errors that were not caught by OCaml's type system whenever I
implemented new features or refactored my code.

\begin{enumerate}

    \item \textbf{Parsing \& Desugaring}. A generated, LR(1) parser parses a text
        file into a syntax tree, which is then desugared into a smaller, more
        concise abstract syntax tree. The former aims to mimic OCaml syntax
        with a few extensions and keywords so that it is familiar and thus easy
        to pick-up for OCaml users. The latter allows for the type-checker to
        be simpler to implement and easier to specify. In general, this part
        will vary for different languages or can be dealt with differently
        using combinators (the EDSL approach) or a syntax-extension if the host
        language offers such support.

    \item \textbf{Type-checking}. The abstract syntax tree is explicitly-typed,
        with some inference to make it less verbose and more convenient to
        write typical programs.

    \item \textbf{Matrix Expressions}. During type-checking, if a matrix-expression
        is encountered, it is either successfully elaborated into an expression
        in the abstract syntax tree which is then consequently type-checked, or
        fails to find suitable routines to calculate the given expression.

    \item \textbf{Code Generation}. The abstract syntax tree is translated into
        standard OCaml and a few-particular `optimisations' are made to produce
        more readable code. This process is type-preserving: the linear type
        system is embedded into OCaml's type system, and so when the OCaml compiler
        compiles the generated code, it acts as a sanity check on the code produced.

    \item \textbf{Executable Artifacts}. A transpiler and a REPL are the main
        artifacts produced for this thesis. For evaluation, I implemented
        Kalman filters in Owl, LT4LA and CBLAS/LAPACKE and a benchmarking
        program to measure execution times.

    \item \textbf{Tests}. As mentioned before, almost all non-trivial functions
        have tests to check their behaviour. The output of the transpiler was
        also tested by having the build system generate OCaml code at compile
        time, which in turn could then be compiled and tested like handwritten
        OCaml code.

\end{enumerate}

\section{Core Language}\label{sec:core_lang}

A full description of the core language can be found in
Appendix~\ref{chap:ott_spec}. For convenience, I have reproduced its type
grammar and defined how I interpret those types in OCaml's type system in
Figure~\ref{fig:type_grammar}.  The point of LT4LA is to keep the same values
and types we are familiar with, but \emph{restrict their usage} in certain
ways. LT4LA's core language's main features are: intuitionistic values,
value-restriction, fractional-capabilities (inferrable at call sites),
if-expressions and recursion.

\begin{figure}[tp]
    \centering
    \begin{minipage}{.3\textwidth}
        \centering
        \begin{grammar}
            <f> ::= `'
            \alt <fc>
            \alt `Z'
            \alt `S' <f>

            <t> ::= `'
            \alt `unit'
            \alt `bool'
            \alt `int'
            \alt `elt'
            \alt <f> `arr'
            \alt <f> `mat'
            \alt `!' <t>
            \alt \synt{$\forall$} <fc.> <t>
            \alt <t> \lit{$\otimes$} \synt{$t'$}
            \alt <t> \lit{$\multimap$} \synt{$t'$}
        \end{grammar}
    \end{minipage}
    \begin{minipage}{.3\textwidth}
        \centering
        \begin{minted}[fontsize=\small, linenos]{ocaml}
module Arr =
  Owl.Dense.Ndarray.D

type z = Z
type 'a s = Succ

type 'a arr =
  A of Arr.arr
  [@@unboxed]

type 'a mat =
  M of Arr.arr
  [@@unboxed]

type 'a bang =
  Many of 'a
  [@@unboxed]
        \end{minted}
    \end{minipage}
    \begin{minipage}{.3\textwidth}
        \begin{align*}
            [\![ f\!c ]\!] &= \texttt{'fc} \\
            [\![ \textbf{Z} ]\!] &= \texttt{z}\\
            [\![ \textbf{S} \, f ]\!] &= [\![ f ]\!]\, \texttt{s}\\
            [\![ \textbf{unit} ]\!] &= \texttt{unit}\\
            [\![ \textbf{bool} ]\!] &= \texttt{bool}\\
            [\![ \textbf{int} ]\!] &= \texttt{int}\\
            [\![ \textbf{elt} ]\!] &= \texttt{float}\\
            [\![ f\, \textbf{arr} ]\!] &= [\![ f ]\!]\, \texttt{arr}\\
            [\![ f\, \textbf{mat} ]\!] &= [\![ f ]\!]\, \texttt{mat}\\
            [\![ \textbf{!} \, t ]\!] &= [\![ t ]\!]\, \texttt{bang}\\
            [\![ \forall f\!c.\, t ]\!] &= [\![ t ]\!]\\
            [\![ t \otimes t' ]\!] &= [\![ t ]\!] \texttt{*} [\![ t' ]\!]\\
            [\![ t \multimap t' ]\!] &= [\![ t ]\!] \rightarrow [\![ t' ]\!]
        \end{align*}
    \end{minipage}
    \caption{Type grammar of LT4LA (left) and an informal interpretation of it
        into OCaml types (right).}\label{fig:type_grammar}
\end{figure}

\subsection{Intuitionistic Values}

To make LT4LA a usable DSL, I needed some way of using values zero or more than
once, as a programmer would an intuitionistic value. For this, I added the
!-constructor at the type-level of the DSL and the \ltfla{Many}-constructor and
the \ltfla{let Many <id> = .. in .. } eliminator at the term-level of the DSL.
The idea behind the !-type is that the value uses no `resources'
(linearly-typed expressions). To start off with, it is enough to say that
anything which can be passed around by copying, will have a !-type. This
includes integers, elements and booleans. So \ltfla{3 : !int} and \ltfla{3. +.
4. : !elt}. However, all bindings are still linear by default, so to
emulate intuitionism, I desugar \ltfla{let !x = <exp> in <body>} to
\ltfla{let Many x = <exp> in let Many x = Many (Many x) in <body>}
(similarly for function argument bindings). The reader can check (using the
rules in Appendix~\ref{chap:ott_spec}) that this has the effect of moving
\ltfla{x : !t} from the linear to the intuitionistic environments, only if
\ltfla{<exp> : !t}.

However, just that desugaring alone is not enough to prevent a user from taking
an array or matrix and moving it into the intuitionistic environments. Why?
There are certain situations in which we \emph{should not} use the \ltfla{Many}
constructor.  Consider the following code: \ltfla{let Many x = Many (array 5)
in <body>}; the expression \ltfla{array 5} uses no linearly-typed variables
from the linear environment. Although we could just reject types of the form
\ltfla{!(_ arr)} to fix this simple example, what about pairs \ltfla{let Many
xy = Many (3, array 5) in <body>}?  Ad-hoc pattern matching on the type cannot
account for all possible situations. With the last case, we can use \ltfla{xy}
as many times as we would like, destruct the pair to get the second component
and thus create \emph{distinct} read-write aliases to the same array.  Alas,
now arrays can be used intuitionistically and all the benefits of linearity are
lost.  Or are they?

\subsection{Value-Restriction}

Not quite, but to understand how we can fix this problem, we need to question
an assumption left implicit up until this point: what does \ltfla{Many} even
mean in the DSL? How is it translated to OCaml? What is the OCaml runtime
behaviour of the DSL's \ltfla{Many}-constructor? One option is to go down the
C++ route and make \ltfla{Many} act like a \ltfla{shared_ptr} and act as a
runtime reference-count for arrays. I chose to not go for this option because
it went against the \emph{explicitness} and \emph{predictability} that C and
Fortran have. It would make analysing when and what is allocated and freed more
like the higher-level languages I was trying to move away from.

My aim is to show linear types are simple to understand and apply to
linear-algebra libraries, enough so that it can be grafted (in a limited way)
on to existing languages as a \emph{library}. In that spirit, the simplest
thing that the DSL's \ltfla{Many}-constructor can mean in the OCaml runtime is
\emph{nothing}.  This language construct is translated into a standard OCaml
type-and-constructor of the form \ltfla{type 'a bang = Many of 'a [@@unboxed]}.
The \emph{unboxed} annotation means that the type and its constructor only
exist for the purpose of type-checking in OCaml; \emph{the OCaml runtime
representation of values of type \ltfla{'a} and \ltfla{'a bang} is exactly the
same}.

With this understanding, our problem is that arrays and matrices are unlike
other values such as integers and elements because (under the OCaml hood)
calling a function with an array argument copies a \emph{pointer} to the array
rather than the array itself, instead of the \emph{value} itself. So, we can
start making a distinction, \emph{defining} elements, integers, booleans,
intuitionistic variables, units and lambda-expressions that capture no linear
variables as \emph{values} (since they cannot break referential-transparency)
and anything else (arrays, matrices, expressions which can be reduced, such as
function application and if-expressions) as not being `values'. If this sounds
familiar, it is because this is the same \emph{value-restriction} `trick' from
the world of polymorphic types applied to linearity instead. We then have the
rule that in LT4LA, we can only use \ltfla{Many} on expressions that are
defined to be \emph{values} and \emph{use no linear variables}.

\subsection{Fractional-Capabilities and Inference}

I started off with linearity and showing how it helps a programmer keep track
of of memory allocation and deallocation, and then demonstrated how to add
intuitionism for the only values we wish to be intuitionistic. I will now
explain how I enforce safe aliasing and read/write permissions.

Array and matrix types are parameterised by
\emph{fractional-capabilities}~\cite{boyland}.  A fraction of 1 ($2^0$)
represents complete ownership of a value; in particular, this allows a
programmer to write or free it. Creating an array gives you ownership of it;
the function \ltfla{array : !int --o z arr} (where \ltfla{z} represents `0').
Once you have ownership of an array, you can free it: \ltfla{free : z arr --o
unit}.  Importantly, because a linear-value may only be used once, the array
just freed is \emph{out of scope} for following expressions, preventing
use-after-free.  Ownership also enables you to write to the array: \ltfla{set :
z arr --o !int --o !elt --o z arr} (the syntax \ltfla{w[i] := j} is just sugar
for \ltfla{set w i j}). Here, linearity prevents accessing aliases which
represented the array \emph{before} the mutation.

Any fraction less than 1 (for simplicity, limited to $2^{-k}$ in this system,
for a positive integer $k$) represents read-only access. So, the \ltfla{'x}
represents a natural number (either a zero \ltfla{z}, variable \ltfla{'x} or a
successor (+1) of a natural number). Hence, you can read from (index) any array
\ltfla{get : 'x .  'x arr --o !int --o 'x arr * !elt} (the syntax \ltfla{let !v
<- w[i]} is just sugar for \ltfla{let (w, !v) = get _ w i}). In general, a
left-arrow \ltfla{<-} signifies transparent rebinding with returned values: it
means a program can \emph{appear} to use a variable multiple times, important
for keeping LT4LA usable and readable. The underscore is how a programmer tells
the compiler to automatically \emph{infer} the correct fractional-capability
based on the other arguments passed to the function. In conjunction with the
requirement that function declarations need type-annotations for their
arguments, \textbf{this allows a fractional-capability to be \emph{correctly
inferred} in \emph{any} LT4LA expression}.

Fractions exist to provide both safe aliasing and read/write permissions, via
the primitives \ltfla{share : 'x . 'x arr --o ('x s) arr * ('x s) arr} and
\ltfla{unshare: 'x .  ('x s) arr --o ('x s) arr --o 'x arr}.  For the former, two
arrays returned (which happen to just be the given array) can now only be read
from and not written to. If you want to write to this array, you must use the
latter to combine other read-only aliases until you are left with a value of
type \ltfla{z arr}, guaranteeing no other aliases exists.

Given this set-up, a programmer now \emph{statically} has \emph{perfect}
information about aliasing and ownership of values in the program. They can
only write to or free an array only when they own it; ownership guarantees no
other aliases exist. In Figure~\ref{fig:ltfla_kalman}, I show how this perfect
information can be used to write more readable (equational and algebraic) code
using easy to reason about, value-semantic expressions whose \emph{intensional}
behaviour (regarding aliasing, read/write permissions and memory allocation,
re-use and de-allocation)  is also explicitly and precisely described and
\emph{automatically checkable}. Now the programmer need not resort to manually
figuring out and inserting \texttt{noalias} annotations and worrying about what
variables can and cannot be written to, re-used or freed; instead they can let
the loyal and tireless compiler do the heavy lifting.

\subsection{If-Expressions}

An if-expression's condition may evaluate either way at run time, so a
programmer must guarantee that both branches use the same set of linear
variables. Writing the type-checker in a pure-functional monadic style paid off
here because I could now sandwich monadic values with state-adjustments either
side of it. Given two monadic values that represented type-checking two
branches of an if-expression, I could use the code in
Figure~\ref{fig:same_resources} to easily save, reset and compare the state
either side of running those monadic values.

\begin{figure}[tp]
    \begin{minted}[fontsize=\small]{ocaml}
let same_resources (wf_a, loc_a) (wf_b, loc_b) =
  let open Let_syntax in
  (* Save state *)
  let%bind {used_vars=prev; env=old_env; _} as state = get in
  (* Reset, run a, save state *)
  let%bind () = put { state with used_vars = empty_used } in
  let%bind res_a = wf_a in
  let%bind {used_vars=used_a; _} as state = get in
  (* Reset, run b, save state *)
  let%bind () = put { state with used_vars = empty_used; env = old_env } in
  let%bind res_b = wf_b in
  let%bind {used_vars=used_b; _} as state = get in
  (* Check if same resources *)
  let keys_a, keys_b = (* convert to (used_a, used_b) to sets *) in
  if Set.equal keys_a keys_b then
    (* merge used_vars and used_b environments *)
  else
    (* report differences *)
    \end{minted}

    \caption{Implementation of same\_resources helper method for type-checking
        if-expressions. Note the monadic style helped compose computations that
        affected the type-checker's state in a simple manner.}\label{fig:same_resources}

\end{figure}

\subsection{Functions and Recursion}

A non-recursive function may be used more than once if it does not refer to any
linear variables from the surrounding scope. So, we can desugar something like
\ltfla{let x = 3 in let !f (y : !int) = x + y in <body>} to \ltfla{let x = 3 in
let Many f =  Many (fun y : !int -> x + y) in  <body>}.  Recursion is slightly
more complicated: we can desugar the factorial function
(Figure~\ref{fig:ltfla_factorial}) to \ltfla{let Many f = fix (f, x : !int, if
(*..*) : !int) in <body> }. However, \ltfla{fix}, like \ltfla{Many}, must also
not use any linear variables from its surrounding scope, because calling a
recursive function may evaluate its body multiple times.

\begin{landscape}
    \begin{figure}[p]
    \centering
    \begin{minipage}{1.47\textheight}
    \inputminted[linenos, fontsize=\footnotesize]{fortran}{kalman.f90}
    \end{minipage}
    \caption{Kalman filter in Fortran 90.}\label{fig:fortran_kalman}
    \end{figure}
\end{landscape}

\section{Matrix Expressions}\label{sec:matrix_exps}

We have now arrived at an extended application of linear types. I will show how
we can apply the ability to automatically check aliasing, read/write
permissions, memory allocation, re-use and deallocation, to the domain of
matrix-expression compilation (automatically generating code like
Figure~\ref{fig:fortran_kalman} from mathematical expressions like
Figure~\ref{fig:maths_kalman}).

In Figure~\ref{fig:fortran_kalman}, we see the difficulty of efficiently
implementing a \emph{Kalman filter}~\cite{kalman}, a powerful set of equations
(Figure~\ref{fig:maths_kalman}) applicable to a wide variety of problems. From
the comments, we see that every variable is annotated with the step/matrix
expression that it will hold at some point during the computation (an
equivalent alternative, say in C++, could be to have a meaningful name for each
step and manually keep track of which names alias the same location).

\begin{figure}[tp]
    \begin{align*}
        \mu' &= \mu + \Sigma H^T (R + H \Sigma H^T)^{-1} (H \mu - \textrm{data})\\
        \Sigma' &= \Sigma ( I - H^T (R + H \Sigma H^T)^{-1} H \Sigma )
    \end{align*}
    \caption{Kalman filter equations (credit:
    \href{http://matthewrocklin.com/blog/work/2012/11/24/Kalman-Filter}{matthewrocklin.com}).}\label{fig:maths_kalman}
\end{figure}

In contrast, Figure~\ref{fig:ltfla_kalman}, offers the advantages of
\begin{itemize}
    \item aliasing: labelling each step with a different, more meaningful variable name,
    \item easily spotting which resources are being passed in and which are
        allocated for the function (new/copy),
    \item unambiguously seeing \emph{when} and what values are freed or written-over;
\end{itemize}
and have the compiler automatically ensure the safety of each of the above by respectively
\begin{itemize}
    \item making it impossible to refer to steps/values which are no longer usable,
    \item ensuring all values are declared and \emph{initialised} correctly before they are used,
    \item checking values are neither used after they are freed/written-over \emph{nor} leaked.
\end{itemize}

Indeed, an inexperienced programmer could take the  na\"ive approach of just
copying sub-expressions by default and then letting the compiler tell it which
copies are never used and removing them systematically until it type-checks.
While it is not quite a black-box, push-button compilation of an expression, I
would argue that, it is just as easy (if not easier) to become familiar with as
Rust and its borrow-checker.

\begin{figure}[tp]
    \inputminted[linenos, fontsize=\small]{ocaml}{../test/examples/kalman.lt}
    \caption{Kalman filter in LT4LA. In contrast with the Fortran
        (Figure~\ref{fig:fortran_kalman}) or C (Figure~\ref{fig:cblas_kalman})
        implementations, this one is readable (equational and algebraic),
        using easy to reason about, value-semantic expressions. In addition to
        that, its \emph{intensional} behaviour(regarding aliasing, read/write
        permissions and memory allocation, re-use and de-allocation) is also
        explicitly and precisely described and (unlike with Fortran or C)
        \emph{automatically checkable}.}\label{fig:ltfla_kalman}
\end{figure}

\subsection{Elaboration}

All of the syntax in Figure~\ref{fig:ltfla_kalman} can be unambiguously
desugared \emph{before} type-checking, through fairly simple pattern matching.
Matrix expression compilation is well-trodden territory in
academia~\cite{rocklin_thesis, fabregat_thesis, gunnels_flame, linnea, taco}
but this is, to my knowledge, the \textbf{first type-based approach} to it.

An overview of the translations are in Figure~\ref{fig:mat_patterns}; details
about choosing between `symm' or `gemm' are omitted for brevity. Before
settling on this approach, I tried implementing a more general type-directed,
nested matrix expression compiler; I will now highlight some of the
difficulties inherent in the problem.

\begin{figure}[tp]
    \begin{minted}[fontsize=\small]{ocaml}
let x <- [| y |] in ..              ==> let (y,x) = copyM_to _ y x in ..
let x <- new [| y |] in ..          ==> let (y,x) = copyM _ y in ..
let x <- [| a*b + c |] in ..        ==> let x <- [| 1.*a*b + 1.*c |] in ..
let x <- [| i*a*b + j*c |] in ..    ==> let ((a,b), c) = (* BLAS *) in ..
let x <- new (m, n) [| a*b |] in .. ==> let c = matrix m n in
                                        let x <- [| 1.*a*b + 0.*c |] in ..
    \end{minted}
    \caption{Syntactic translations of matrix expressions to linearly-typed
        matrix functions. Further annotations on the matrix variables
        (\ltfla{sym} or \ltfla{^T}) determine which BLAS routine is called and
        what parameters it is passed.}\label{fig:mat_patterns}

\end{figure}

One of the first hurdles I encountered was compositionality: to compile $AB +
C$ it is typically better to use a BLAS routine directly rather than but first
compiling $A$, then $B$, then adding a call to multiply them, then compiling
$C$ and finishing with a call to add the results.

Another compositionality problem is that a call to a linear function does not
just return a result, but a sequence of re-bindings that dictate which
variables are still in scope/usable.  As such, to compile an expression, you
need to provide a CPS-style function of type $var \rightarrow exp$ representing
how you would use a variable representing the result in the rest of the
expression.

However, the type of that variable also determines how you can use it: can you
write to it or must you copy it? This information depends on how the expression
representing the variable was elaborated and adds to the complexity of the
pattern-matching.

Copying leads us into dealing with temporaries: do you first allocate all
temporaries in new matrices (SSA-style) and then analyse dimensions to figure
out which slots of memory can be re-used via copy coalescing? Or do you try and
infer the live ranges of available resources from the environment as you go?
Can, and should, you type arrays and matrices (or $n$-dimensional tensors) as
the same?

Adding in more and more considerations, the problem starts to resemble register
allocation: there are registers of different types and sizes, many
(non-orthogonal) instructions to choose from, a cost model to take into account
all whilst trying to balance the number of registers in use and the number of
instructions emitted.

\section{Code Generation}

Code generation is a straightforward mapping from core LT4LA constructs to
OCaml constructs, with the addition of \ltfla{Many} constructors to wrap
integer, element and boolean literals.

To make the code produced readable, I added a few `optimisations', to the
compiler, to `re-sugar' some of the constructs translated where appropriate.
Almost all of them involved the erasing the !-eliminator which is not needed in
regular, intuitionistic OCaml. So, so we can simply replace any expression of
the form:

\begin{itemize}

    \item \ltfla{let xy, z = <exp> in let x,y = xy in <body>} with
        \ltfla{let (x,y), z = <exp in <body>},

    \item \ltfla{let Many x = x in let Many x = Many (Many x) in <body>} with
        \ltfla{<body>},

    \item \ltfla{let Many x = <exp> in let Many x = Many (Many x) in <body>} with
        \ltfla{let x = <exp> in <body>},

    \item \ltfla{let Many x = Many <exp> in <body>} with \ltfla{let x = <exp> in <body>},

    \item \ltfla{let Many f = fix (f, x, .., <exp>, ..) in <body>} with
        \ltfla{let rec f x = <exp> in <body>}.

\end{itemize}

The end result is visible in Figures~\ref{fig:ocaml_sumarray}
and~\ref{fig:ocaml_kalman}:  (surprisingly beautiful) OCaml code that is not
obviously safe, as promised at the start of the chapter.

\begin{figure}[tp]
    \centering
    \begin{minted}[linenos, fontsize=\small]{ocaml}
let rec f i n x0 row =
  if Prim.extract @@ Prim.eqI i n then (row, x0)
  else
    let row, x1 = Prim.get row i in
    f (Prim.addI i (Many 1)) n (Prim.addE x0 x1) row
in
f
    \end{minted}
    \caption{Recursive OCaml function for a summing an array, generated (at
        \emph{compile time}) from the code in Figure~\ref{fig:ltfla_sumarray},
        passed through \texttt{ocamlformat} for presentation.}\label{fig:ocaml_sumarray}

\end{figure}

It is clear that both Figures~\ref{fig:fortran_kalman}
and~\ref{fig:ocaml_kalman} are realisations of a concise, linearly-typed Kalman
filter~\ref{fig:ltfla_kalman} \emph{specification} that describes the whole
behaviour and intent (with respect to read/writes permissions, memory
management and aliasing) of the program and the BLAS primitives it uses more
accurately than a Fortran, C or OCaml implementation could.

\begin{figure}[p]
    \begin{minted}[linenos, fontsize=\footnotesize]{ocaml}
let kalman sigma h mu r_1 data_1 =
  let h, _p_k_n_p_ = Prim.size_mat h in
  let k, n = _p_k_n_p_ in
  let sigma_h = Prim.matrix k n in
  let (sigma, h), sigma_h =
    Prim.symm (Many true) (Many 1.) sigma h (Many 0.) sigma_h
  in
  let (sigma_h, h), r_2 =
    Prim.gemm (Many 1.) (sigma_h, Many false) (h, Many true) (Many 1.) r_1
  in
  let (h, mu), data_2 =
    Prim.gemm (Many 1.) (h, Many false) (mu, Many false) (Many (-1.)) data_1
  in
  let h, new_h = Prim.copy_mat_to h sigma_h in
  let r_2, new_r = Prim.copy_mat r_2 in
  let chol_r, sol_h = Prim.posv new_r new_h in
  let chol_r, sol_data = Prim.potrs chol_r data_2 in
  let () = Prim.free_mat chol_r in
  let h_sol_h = Prim.matrix n n in
  let (h, sol_h), h_sol_h =
    Prim.gemm (Many 1.) (h, Many true) (sol_h, Many false) (Many 0.) h_sol_h
  in
  let () = Prim.free_mat sol_h in
  let h_sol_data = Prim.matrix n (Many 1) in
  let (h, sol_data), h_sol_data =
    Prim.gemm (Many 1.) (h, Many true) (sol_data, Many false) (Many 0.) h_sol_data
  in
  let mu, mu_copy = Prim.copy_mat mu in
  let (sigma, h_sol_data), new_mu =
    Prim.symm (Many false) (Many 1.) sigma h_sol_data (Many 1.) mu_copy
  in
  let () = Prim.free_mat h_sol_data in
  let h_sol_h_sigma = Prim.matrix n n in
  let (sigma, h_sol_h), h_sol_h_sigma =
    Prim.symm (Many true) (Many 1.) sigma h_sol_h (Many 0.) h_sol_h_sigma
  in
  let sigma, sigma_copy = Prim.copy_mat_to sigma h_sol_h in
  let (sigma, h_sol_h_sigma), new_sigma =
    Prim.symm (Many false) (Many (-1.)) sigma h_sol_h_sigma (Many 1.) sigma_copy
  in
  let () = Prim.free_mat h_sol_h_sigma in
  ((sigma, (h, (mu, (r_2, sol_data)))), (new_mu, new_sigma)) )
in
kalman
    \end{minted}
    \caption{OCaml code for a Kalman filter, generated (at \emph{compile time})
        from the code in Figure~\ref{fig:ltfla_kalman}, passed through
        \texttt{ocamlformat} for presentation.}\label{fig:ocaml_kalman}

\end{figure}

\subsection{Build System}

So once you have written your memory-optimised program with all the features
and support provided by LT4LA and made it produce well-typed, compilable OCaml
code, the question then becomes, how to use this code. This process will vary
across ecosystems, but within OCaml, the new build system on the block
\emph{Dune} has support for generating, compiling and linking OCaml modules at
\emph{compile time} (Figure~\ref{fig:build}). Indeed, this is how I have written
tests and benchmarks for programs produced by LT4LA from within OCaml. In
particular, I can use my linearly-typed Kalman filter implementation just like
and with any other OCaml function.

\begin{figure}[t]
    \centering
    \includegraphics[width=\textwidth]{impl_build}
    \caption{Using LT4LA functions from OCaml with the Dune
        build system.}\label{fig:build}
\end{figure}

Advantages of generating code at compile time include avoiding runtime overhead
as well as catching any interface changes (type-signatures now include
\emph{intensional} information about which parameters are read from or written
to) between the generated code and the code that uses it via OCaml's type
system, \emph{another benefit} (in addition to a sanity check on LT4LA's
type-checker implementation) of \emph{embedding} LT4LA's type system inside of
OCaml's.

I suspect that this approach will be very valuable to not just \emph{users} of
libraries such as Numpy or Owl, but also their \emph{implementors}: library
functions which present a safe, value-semantic interface but use unsafe,
mutating operations on the inside could now be expressed using LT4LA and have
an extra safety-net and automatic checking for their implementations.

\section{Summary}

I have explained how a few core features -- linearity, the \ltfla{Many}
constructor, value-restriction, fractional-capabilities with inference,
if-expressions and recursive functions -- are enough to \emph{statically
capture and automatically check} aliasing, read/write permissions, memory
allocation, re-use and deallocation of non-trivial linear-algebra programs. I
have also demonstrated that simple pattern-matching and desugaring provides the
potential for a new, \emph{type-directed} approach to matrix expression
compilation. Lastly I have shown that it is possible to use these features with
\emph{existing} languages and frameworks.\footnote{As mentioned in the previous
chapter, if the host language supports \emph{syntax-extensions}, like PPX
for OCaml, it is possible to construct LT4LA expressions \emph{from within} the
host language.}