### 2018-05-27 write-up/introduction.tex
Finish Introduction

\chapter{Introduction}

% No resetting page numbers: ISO 7144
% \pagenumbering{arabic}
% \setcounter{page}{0}

% \begin{guidance}
%     This is the introduction where you should introduce your work.  In
%     general the thing to aim for here is to describe a little bit of the
%     context for your work --- why did you do it (motivation), what was the
%     hoped-for outcome (aims) --- as well as trying to give a brief
%     overview of what you actually did.
% 
%     It's often useful to bring forward some ``highlights'' into
%     this chapter (e.g.\ some particularly compelling results, or
%     a particularly interesting finding).
% 
%     It's also traditional to give an outline of the rest of the
%     document, although without care this can appear formulaic
%     and tedious. Your call.
% \end{guidance}

Linear types allow the compiler and programmer to statically keep track of the
resources that an implementation uses, thus offering a promising solution to
resource management. However, they have not made their way into mainstream
programming languages. To demonstrate their usefulness, I implemented an OCaml
library that allows users to learn about, use and become familiar with linear
types, specifically in the context of linear-algebra programs.

The main contributions of this thesis are:

\begin{itemize}

    \item An \textbf{original, usable implementation} of a type system that can express
        aliasing, read/write permissions, memory allocation, re-use and deallocation.

    \item An \textbf{original} demonstration of how that type system can be
        \textbf{applied to the APIs} of linear algebra libraries and the
        \textbf{benefits} of doing so.

    \item Many \textbf{new} examples of how the type system can \textbf{automatically
        check} for common aliasing, read/write permissions, memory allocation,
        re-use and deallocation \textbf{errors} in the context of
        linear-algebra programs.

    \item An \textbf{original} demonstration of how that type system can be
        \textbf{useful} for \textbf{matrix expression compilation}.

    \item \textbf{New and readable implementations} of \textbf{non-trivial}
        linear-algebra programs that \textbf{take advantage} of said type
        system.

    \item A \textbf{new solution} to the \textbf{dichotomy} of
        \textbf{readability}, \textbf{ease of reasoning and safety} of
        high-level linear-algebra libraries versus the
        \textbf{memory-efficiency} of low-level linear-algebra libraries.

    \item A \textbf{new library design} to provide said solution in a way that
        \textbf{integrates well} with existing OCaml code and linear-algebra
        libraries.

\end{itemize}