### 2018-05-28 write-up/background.tex

\chapter{Background}

% \begin{guidance}
%     A more extensive coverage of what's required to understand your
%     work. In general you should assume the reader has a good undergraduate
%     degree in computer science, but is not necessarily an expert in
%     the particular area you've been working on. Hence this chapter
%     may need to summarize some ``text book'' material.
%
%     This is not something you'd normally require in an academic paper,
%     and it may not be appropriate for your particular circumstances.
%     Indeed, in some cases it's possible to cover all of the ``background''
%     material either in the introduction or at appropriate places in
%     the rest of the dissertation.
% \end{guidance}

\prechapter{%
    I will outline the concept of linear types and show how they can be used to
    solve the problems faced by programmers writing code that uses linear algebra
    libraries. I will emphasise \emph{practical} and intuitive explanations of
    linear types to keep this thesis accessible to programmers as well as
    academics not familiar with type theory and will give only a terse overview
    of the history and theory behind linear types for the interested reader to
    pursue further.
}%

\section{Tracking Resources with Linearity} Familiar examples of using a type
system to express program invariants are existential types for abstraction and
encapsulation, and polymorphic types for parametricity and composition (a.k.a
generics). Less-known examples include dependent types for contracts or pre-
and post-conditions. The advantages of using a type system to express program
invariants are summarised by saying: the stronger the rules you follow, the
better the guarantees you get about your program \emph{before} you run it. At
first, the rules may seem restrictive, but similar to how the rules of grammar,
spelling and more generally writing help a writer communicate their ideas more
clearly, so too do typing rules make it easier for the programmer to
communicate the intent and assumptions of their programs. An added, but often
overlooked benefit is automated-checking: a programmer can boldly refactor
their programs and the type checker will \emph{assist} in enforcing the
invariants it was given by highlighting where they were violated.

Linear types are a way to help a programmer track and manage resources. In
practical programming terms, they enforce the restriction that a value may be
used exactly once\footnote{This definition may differ from more colloquial uses
in discussions surrounding \emph{substructural} type systems and/or Rust.}
(Figure~\ref{fig:example_rules} shows two example typing rules for a typical,
linearly typed language). While this restriction may seem limiting at first,
precisely this constraint can be used to express common invariants of typical
programs. For example: a file or a socket, once opened \emph{must} closed; all
memory that is manually allocated \emph{must} be freed.  C++'s destructors and
Rust's Drop-trait (and more generally, its borrow-checker) attempt to enforce
these constraints by basically doing the same thing: any resource that has not
been moved is deallocated at the end of the current lexical scope. Notably,
these languages also permit aliasing, alongside rules enforcing when it is
acceptable to do so.  On face value, the above one-line description of linear
types prevents aliasing or functions such as $\lambda x.\; (x, x)$; such
features are still allowed (albeit in a more restricted fashion) in a
\emph{usable} linear type system designed for working with linear algebra
libraries.

\begin{figure}
    \centering
    \begin{minipage}{0.3\textwidth}
        \centering
    \begin{prooftree}
        \AxiomC{}
        \RightLabel{\textsc{Var}}
        \UnaryInfC{$\Gamma, x : t \vdash x : t ; \Gamma$}
    \end{prooftree}
    \end{minipage}%
    \begin{minipage}{0.6\textwidth}
        \centering
    \begin{prooftree}
        \AxiomC{$\Gamma_1 \vdash e_1 : t_1 ; \Gamma_2$}
        \AxiomC{$\Gamma_2 \vdash e_2 : t_2 ; \Gamma_3$}
        \RightLabel{\textsc{Pair}}
        \BinaryInfC{$\Gamma_1 \vdash (e_1, e_2) : t_1 \otimes t_2 ; \Gamma_3$}
    \end{prooftree}
    \end{minipage}
    \caption{A typing rule has the general form $\Gamma_{\textrm{in}} \vdash e ; \Gamma_{\textrm{out}}$.
    Typing a variable \emph{removes} it from the environment. Typing a pair
    requires that each component of the pair uses \emph{different} resources
    from the environment.}\label{fig:example_rules}
\end{figure}

\section{Problem in Detail}

Given this background, the most pertinent question at hand is: what problems do
linear algebra library users (and implementors) typically face? The answer to
this question depends on which of two buckets a programmer falls (or is forced
by domain) into. On one side, we have users of high-level frameworks such as
Owl (for OCaml), Julia and Numpy (for Python); on the other, we have users
of more manual, low-level libraries such as BLAS (Basic Linear Algebra
Subroutines) for languages like C, C++ and Fortran.\footnote{I am not including
Rust in this comparison because its linear algebra libraries are under
active development and not as well-known/used. Later on, I will compare and
contrast Rust (a language with in-built support of substructural features
to track resources) with this project to evaluate the classic
(E)DSL-versus-language-feature debate as it applies to linear types.} Most of
what follows applies to \emph{dense} linear algebra computations rather than
\emph{sparse} ones because for the latter, memory allocated for results depends
on the sparsity of the inputs and so is not immediately amenable to the
techniques proposed in this thesis.

\subsection{One Too Many Copies and a Thousand Bytes Behind}
\begin{figure}[tbp]
    \centering
    \begin{minipage}{0.45\textwidth}
    \centering
    \begin{minted}[fontsize=\small]{python}
# Numpy (Python)
import numpy.matlib
a = [[1,0],[0,1]]
b = [[4,1],[2,2]]
c = numpy.matmul(a,b)
    \end{minted}
    \begin{minted}[fontsize=\small]{julia}
# Julia
c = [1 0; 0 1] * [4 1; 2 2]
    \end{minted}
    \end{minipage}%
    \begin{minipage}{0.45\textwidth}
    \centering
    \begin{minted}[fontsize=\small]{ocaml}
(* Owl (OCaml) *)
open Owl
let a = Mat.of_arrays
  [|[|1.;0.|];[|0.;1.|]|]
let b = Mat.of_arrays
  [|[|4.;1.|];[|1.;2.|]|]
let c = Mat.(a *@ b)
    \end{minted}
    \end{minipage}
    \caption{Matrix Multiplication in Numpy (Python), Julia and Owl (OCaml).}\label{fig:mat_mul_copies}
\end{figure}

In Figure~\ref{fig:mat_mul_copies}, we see that matrix multiplication is fairly
trivial to write and execute in Numpy, Julia and Owl. Let us call this approach
\emph{value-semantic}, meaning that objects are \emph{values} just like
integers and floating-point numbers. This approach confers two key advantages to the
programmmer: it is easy to read (equational and algebraic expressions) and it
is easy to reason about (as one would with a mathematical formula). Although this
approach does permit \emph{aliasing}, the consequences are benign because the
result of any computation is a \emph{new} value, distinct from any used during
the calculation of that value.

\begin{figure}[tbp]
    \centering
    \begin{minted}[fontsize=\small]{ocaml}
let mul x y =
  if same_shape x y then
    let y = copy y in
    (_owl_mul (kind x) (numel x) x y y; y)
  else
    broadcast_op (_owl_broadcast_mul (kind x)) x y
    \end{minted}
    \caption{Implementation of Matrix Multiplication in Owl (OCaml).
        Note the `copy' for the result and the unsafe `\_owl\_mul' operation
    used to perform an in-place multiplication.}\label{fig:mat_mul_owl}
\end{figure}

However, these advantages come with some costs: constantly producing new values
is wasteful on memory (although the example given in
Figure~\ref{fig:mat_mul_copies} is only a $2 \times 2$ matrix, many real-world
datasets can contain up to gigabytes of data). A complex expression may create
many short-lived temporaries which would need to be reclaimed by a
garbage-collector (see Figure~\ref{fig:mat_mul_owl}). Libraries taking a
\emph{value-semantic} approach offer a dichotomy for a user wishing to
implement a new algorithm: either use the existing and safe primitives to build
an easy to reason about but slower, more memory-intensive algorithm, or use
escape-hatches (typically provided by most libraries, which permit in-place
modification of objects) to build faster, less readable and more
memory-efficient algorithms which are harder to reason about.

\subsection{IHNIWTLM}

The title of this subsection\footnote{I Have No Idea What Those Letters Mean.}
illustrates the problem with low-level libraries: readability (equational and
algebraic expressions) and ease of reasoning is sacrificed at the altar of
performance and (memory) efficiency.

Although escape-hatches do exist in high-level libraries, their use is
discouraged. Systematic consideration of performance requires lowering the
level of abstraction a programmer is working on. At this level, \textbf{several}
factors such as memory layout, allocation, re-use as well as cache behaviour
and parallelism are important. Of these, \textbf{only memory allocation and
re-use} are relevant to linear types and this thesis.

\begin{figure}[tbp]
    \begin{minted}[fontsize=\small]{fortran}
  program blasMatMul
  implicit none
  real*4 a(2,2), b(2,2), c(2,2)
C External from BLAS
  external dgemm
C Initialize in column major storage of Fortran
  data a/ 1,0,0,1/
  data b/ 4,1,1,2/
C            tfm  tfm  rowA colB K  alpha a lda  b  ldb beta c  ldc
  call dgemm('N', 'N', 2,   2,   2, 1.0,  a,  2, b, 2,  0.0, c, 2)
    \end{minted}
    \caption{One of \emph{several} BLAS (Fortran) routines for Matrix
    Multiplication.}\label{fig:fortran_blas}
\end{figure}

In Fortran (Figure~\ref{fig:fortran_blas}) and C, temporary storage for all
intermediate values must be managed by the programmer. While this approach
leads to verbose, less readable and harder to reason about code, the
explicitness is good for understanding the memory concerns of the program.

On the other hand, C++ (with operator overloading) can end up looking fairly
readable. For safety and correctness, expressions are typically handled with
value-semantics (the result of any computation is a \emph{new} value, distinct
from any used during the calculation of that value). However, given
\emph{extra} information about aliasing (Eigen, Figure~\ref{fig:cpp_eigen}) or
usage of intermediate expressions (uBLAS, Figure~\ref{fig:cpp_ublas}), the
number of temporaries allocated can be reduced and increased \emph{implicitly}
to improve performance (remove unnecessary allocations or re-calculations
respectively). Further tricks to improve performance include expression
templates (building up an expression-tree at compile time and then
pattern-matching on it to produce code) and lazy evaluation (only calculating a
result when it is needed). These will be discussed in more detail in
Chapter~\ref{chap:related}.

It is important to note that should a \texttt{noalias} annotation be wrong, the
program's behaviour is very likely to end up being undefined (like how
\texttt{memcpy()} for overlapping regions of memory is explicitly undefined in
the POSIX and C standards). Indeed, one of Fortran's strengths lies in assuming
that references cannot be aliased (with certain caveats) in more cases than C
permits (this informal, general statement comes with many nuances left for the
interested reader to pursue).

\begin{figure}[tp]
    \begin{minted}[fontsize=\small]{c++}
#include <iostream>
#include <Eigen/Dense>
using namespace std;
int main()
{
    Eigen::Matrix2d a,b,c;
    a << 1, 0, 0, 1; b << 4, 1, 1, 2; c << 0, 0, 0, 0;
    a * b; // new matrix
    c += a * b; // temporary for correctness in case of aliasing
    c.noalias() += a * b; // no temporaries
}
    \end{minted}
    \caption{Some examples of Matrix Multiplication in Eigen. Using expression
        templates (to be discussed later) and \emph{explicitly provided} aliasing
        information, Eigen can emit a single BLAS `dgemm'-like call for the last
	line, mirroring the Fortran example of
	Figure~\ref{fig:fortran_blas}.}\label{fig:cpp_eigen}
\end{figure}

\begin{figure}[tbp]
    \begin{minted}[fontsize=\small]{c++}
noalias(C) = prod(A, B);
// Preferable if T is preallocated
temp_type T = prod(B,C); R = prod(A,T);
prod(A, temp_type(prod(B,C));
prod(A, prod<temp_type>(B,C));
    \end{minted}
    \caption{Boost uBLAS example of Matrix Multiplication. Temporaries need to
        be marked as such to prevent unnecessary re-computation of
        values.}\label{fig:cpp_ublas}
\end{figure}

\section{Proposed Solution}\label{sec:prop_sol}

My proposed solution to this dichotomy (readability, ease of reasoning and
safety versus memory-efficiency) is a \emph{domain-specific language} (DSL),
called LT4LA (Linear Types for Linear Algebra). It is written in OCaml and
transpiles to OCaml. It offers readable, explicit management of aliasing,
read/write permissions, memory allocation, re-use and deallocation all with
automated checking: offering a safety net to catch the baby whilst swiftly
disposing of the bath water.  Although for expository and testing purposes I
have defined a concrete-syntax, a full implementation would make use of a
language's syntax-extension features (such as PPX for OCaml) to \emph{embed}
the DSL into the host language. Such an embedding is straightforward but fairly
tedious to implement. As a half-way point, I used compile-time code generation
to make the DSL's output available to OCaml for testing and evaluation.

Let us have a look at a few examples of functions we can write with linear
types.  We can define the factorial function
(Figure~\ref{fig:ltfla_factorial}) and sum over an array
(Figure~\ref{fig:ltfla_sumarray}).

\begin{figure}[tp]
    \inputminted[linenos, fontsize=\small]{ocaml}{../test/examples/factorial.lt}
    \caption{Factorial function in LT4LA.}\label{fig:ltfla_factorial}
\end{figure}

\begin{figure}[tp]
    \inputminted[linenos, fontsize=\small]{ocaml}{../test/examples/sum_array.lt}
    \caption{Summing over an array in LT4LA.}\label{fig:ltfla_sumarray}
\end{figure}

The syntax is intended to resemble OCaml's, apart from the spurious
\ltfla{`!'}s found here and there (they are annotations to show that we can
use a value more than once). In Figure~\ref{fig:ltfla_sumarray}, we see
\ltfla{row} has type (\ltfla{'x arr}). More detailed explanations of
\emph{what and why} will be given in Chapter~\ref{chap:impl}, but for now, it
is enough to know it means we can only \emph{read} from \ltfla{row}, and not
write to it. If we did try to write to or free \ltfla{row}, we would get a
helpful error message, as shown in Figure~\ref{fig:ltfla_univ}.

\begin{figure}[tp]
    \begin{minted}[linenos, firstnumber=7, fontsize=\small]{ocaml}
let row = row[i] := x1 in (* or *) let () = free row in
(* Could not show equality:                                        *)
(*     z arr                                                       *)
(* with                                                            *)
(*     'x arr                                                      *)
(*                                                                 *)
(* Var 'x is universally quantified                                *)
(* Are you trying to write to/free/unshare an array you don't own? *)
(* In test/examples/sum_array.lt, at line: 7 and column: 19        *)
    \end{minted}
    \caption{Attempting to write to or free a read only array in
    LT4LA.}\label{fig:ltfla_univ}
\end{figure}

Now suppose we are trying to square a matrix, using a `dgemm' like BLAS
routine (called `simple\_dgemm'\footnote{For the purposes of this example, I
assume `simple\_dgemm' has type \ltfla{'x . 'x mat --o 'y .  'y mat --o z mat
--o ('x mat * 'y mat) * z mat}.} in Figures~\ref{fig:ltfla_square}
and~\ref{fig:ltfla_free}) which takes two read-only matrices and a third matrix
it can write to and performs $C := AB + C$. How would we use such a routine to
square a matrix, assuming $C=0$ initially?  Surely this would break linearity,
since $A$ would have to be passed in to the function twice, and we can only 
use non-!-annotated variables once?

To solve this, we use a special primitive called \ltfla{share} to produce
(more) read-only \emph{aliases} of any matrix. We then pass these into the
function, and it works as expected. Once the squared matrix has been obtained,
we may not want the original anymore, and thus decide to free it. Before we do
so, we must first \ltfla{unshare} any read-only aliases that exist, to get back
a single, read-write handle.

\begin{figure}[tp]
    \begin{minted}[linenos, fontsize=\small]{ocaml}
let (a1, a2) = share _ a in
let ((a1, a2), c) = simple_dgemm _ a1 _ a2 c in
let a = unshare _ a1 a2 in
let () = free a in c
    \end{minted}
    \caption{Squaring a matrix in LT4LA (assuming \ltfla{c} is initially
        0).}\label{fig:ltfla_square}
\end{figure}

If we tried to free one of the read-only aliases before or instead of
\ltfla{unshare}-ing them, then we would get the error shown in
Figure~\ref{fig:ltfla_free}. Briefly, a \ltfla{z arr} is a read-write array,
everything else (\ltfla{'x arr} or \ltfla{_ s arr}) is read-only. The types of
\ltfla{free}, \ltfla{share} and \ltfla{unshare} (see
Appendix~\ref{chap:primitives}) are set up so that you can only free something
when you have read-write access to it, guaranteed by linearity to be the only
name in scope with this capability (aliases can only be read-only). Conversely,
if we \emph{forgot} to \ltfla{free a}, we would also get \texttt{Variable a not
used} error.

Another way in which `simple\_dgemm' may be misused is by instantiating its
informal description of $C := AB + C$ with $B = C = A$ and so mistakenly
concluding that it computes $A := A^2 + A$ in-place. However, the type of
\ltfla{share} prevents this as well -- \ltfla{let (a11, a12) = share _ a1 in
simple_dgemm _ a11 _ a12 a2} would result in an error similar to the one shown
in Figure~\ref{fig:ltfla_free}.

\begin{figure}[tp]
    \begin{minted}[linenos, fontsize=\small]{ocaml}
let (a1, a2) = share _ a in
let ((a1, a2), c) = simple_dgemm _ a1 _ a2 c in
let () = free a1 in c
(* Error:                                                        *)
(* Could not show equality:                                      *)
(*     z arr                                                     *)
(* with                                                          *)
(*     z s arr                                                   *)
(*                                                               *)
(* Could not show z and z s are equal.                           *)
(* Are you trying to write to/free an array before unsharing it? *)
(* In test.lt, at line: 3 and column: 17                         *)
    \end{minted}
    \caption{Attempting to free a read-only alias of matrix.}\label{fig:ltfla_free}
\end{figure}

\section{Further Reading and Theory}\label{sec:further}

% Just as no Scottish MP's initial speech in the House of Commons would be
% complete without some claim or mention to Rabbie Burns,

No exposition of linear types would be complete without a mention of Girard's
Linear Logic~\cite{girard}. As mentioned in the Stanford Encyclopedia of
Philosophy, it is ``a refinement of classical and intuitionistic logic. Instead
of emphasizing truth, as in classical logic, or proof, as in intuitionistic
logic, linear logic emphasizes the role of formulas as resources.'' A walk from
logic to programming along the well-trodden Curry-Howard bridge brings us to
linear types~\cite{wadler}.

For the category theory inclined reader, the !-operator (sometimes, for reasons
elided here, called \emph{exponentiation}) forms a co-monad; for the rest of
us, this entails two (rather simple) facts about a value you can use any number
of times: you can (1) can use it once (co-unit), and (2) can pass it to many
contexts that will use it many times (co-multiply).

More generally, by annotating variables in the context with their usage (when
implementing a type checker for a linearly typed language), we can express the
rules of \emph{substructural} (including affine, relevant and ordered type
systems) under the more general framework of \emph{co-effects}~\cite{petricek}.

Stepping further back, both the practice and theory behind resource-aware
programming has made visible progress in the past few years.  On the
programming side, we have Linear Haskell, Rust and Idris (experimental). On the
research side, we have Resource Aware ML~\cite{hoffmann} and the tantalising
promise of integrating linear and dependent types~\cite{atkey}.

\section{Summary}

I have given an \emph{intuitive} exposition of linear types with
\emph{fractional-capabilities}~\cite{boyland}, emphasising small, but
illustrative and practical code examples, leaving \emph{what} is going on and
\emph{why} it works as details for the next chapter. I have shown that it is
possible to solve the dichotomy of readability, ease of reasoning and safety
versus memory-efficiency by providing explicit and automatically checked
management of aliasing, read/write permissions, memory allocation, re-use and
deallocation with a \emph{type system}. This prevents a \textbf{whole class} of
errors that can occur with low-level languages \emph{at compile time}. For
example, \textbf{the types make it impossible to use `simple\_dgemm'
incorrectly}. I demonstrated these features using the context of linear algebra
libraries as as a specific example. In the next chapter, I will show that these
features can be further applied to the problem of \emph{matrix expression
compilation}. I will also explain how to express and implement them so that
they can be provided and used \emph{as a library}.