### 2018-05-28 write-up/conclusion.tex

\chapter{Conclusion}

% \begin{guidance}
%     As you might imagine: summarizes the dissertation, and draws
%     any conclusions. Depending on the length of your work, and
%     how well you write, you may not need a summary here.
%
%     You will generally want to draw some conclusions, and point
%     to potential future work.
% \end{guidance}

In this thesis, I presented \textbf{linear types with fractional-capabilities}
as a type-based formalism for expressing \textbf{aliasing, read/write
permissions, memory allocation, re-use and deallocation}; I provided a detailed
description of each feature of the type system, its intent and implementation
in Section~\ref{sec:core_lang}.

I used that formalism to make \textbf{precise and explicit the intensional
behaviour and assumptions} (about aliasing, read/write permissions and memory
re-use) of linear-algebra library APIs, first with the simplified example of
the fictional `simple\_dgemm' primitive in Section~\ref{sec:prop_sol} -- where
I showed how \textbf{the types made it impossible to use `simple\_dgemm'
incorrectly} -- and then, more realistically with the types of LT4LA primitives
(including actual CBLAS/LAPACKE routines) listed in
Appendix~\ref{chap:primitives}.

I also used that formalism to implement a simple, yet effective matrix
expression `compiler' based on pattern-matching. I was able to use to that
compiler to write, a \textbf{non-trivial, yet readable} (equational and
algebraic) linear algebra program (Kalman filter, as described in
Section~\ref{sec:matrix_exps}). I shared how, in doing so, I could use the
type-checking as a tool to \textbf{automatically highlight bugs} regarding
aliasing, read/write permissions and memory re-use of a reference
implementation (Section~\ref{sec:qual_benefits}) and the matrix expression
compiler that generated it (Section~\ref{sec:mat_exp_comp}).

% This suggests that although it is tempting to think of matrix expression
% compilation as a data-flow problem in terms of graphs, such an approach may
% struggle to accurately capture the intensional behaviour and assumptions of
% linear-algebra kernels which seem to be expressed nicely with linear types and
% fractional-capabilities.

I showed that this implementation was \textbf{more memory-efficient} than one
written using a high-level library and performed just as
\textbf{predictably} as one written using a low-level library,
enabling a programmer to have the best of both worlds
(Chapter~\ref{chap:eval}). The type system ensured that the implementation was
\textbf{safe and explicit} with respect to aliasing, read/write permissions,
memory allocation, re-use and deallocation.

Lastly, I was able to provide all of this functionality as a \textbf{usable
OCaml library}, by generating \textbf{readable and not obviously safe} OCaml
code (1) which looks like it is written by an expert and (2) whose types
reflect its intensional behaviour and assumptions (interpreting types into
OCaml, Figure~\ref{fig:type_grammar}). This library, LT4LA, produces
\textbf{helpful type-errors} (Section~\ref{sec:prop_sol}), comes with an
interactive REPL and is documented and well-tested
(Section~\ref{sec:structure_lt4la}).

\section{Future Work}

LT4LA is a proof-of-concept design for linearly typed DSL for writing
linear-algebra programs in a way that can work with existing languages and
libraries. It is made of up of the core language and its matrix expression
`compiler'; both could be extended in many ways.

\subsection{Core Language}

Although I use well-established type system features, it may be worthwhile to
formally state and verify my claimed safety properties.  The logical conclusion
of LT4LA's core language would be a language expressive enough to provide a
usable, linearly typed interface to the unsafe parts of the Owl library for
OCaml, perhaps as a functor like its current, `Lazy' interface.  To do that,
the language implementation would need to support more than just 64-bit
floating-point numbers for elements (for example, 16-bit, 32-bit or complex),
polymorphism, a unified way of dealing with arrays and matrices (perhaps
through sub-typing). It may also be useful to explore localised linearity
\emph{inference}, to automate the reasoning I did manually when implementing a
memory-efficient Kalman filter (Section~\ref{sec:qual_benefits}).

Another, equally valid but opposing direction may be to decrease the amount of
expressivity, to bring it closer to Fortran and C so that it would be easier to
emit code in those languages. A full set of linearly typed BLAS/LAPACK bindings
(to encode information in documentation at the type level) would be a good
first step, perhaps leading into more research on formal semantics for Fortran.

\subsection{Matrix Expression Compilation}

All of the matrix expression compilers mentioned in the previous chapter
construct some sort of data-flow graph to represent the computation being
executed. While this seems intuitive, there is no formal argument for this
approach.  Some directions in which a type-based approach to efficient matrix
expression compilation could be taken are:

\begin{itemize}

    \item \textbf{As a typed IR} for matrix expression compilers. This in turn
        could enable
        \begin{itemize}
            \item existing matrix expression compilers to be less opaque about
                what resources they are consuming.
            \item open up opportunities for non-local sharing of temporary values
                with some intra-procedural analysis.
            \item allow the user to choose: use a matrix expression compiler when
                desired and drop down to a usable typed IR for finer control,
                whilst still retaining safety guarantees.
        \end{itemize}

    \item \textbf{Formal verification of matrix expression compilers} by
        precisely specifying the intensional and extensional properties of the
        source and target languages.

    \item \textbf{Multi-stage programming} to use information only available at
        runtime (such as sizes, matrix properties, sparsity, control flow) for
        generating more specialised code.

    \item \textbf{Dependent types} to have control over how resources can be
        used and split. In addition to formal verification, dependent types
        could be combined with linearity to express finer-grain conventions
        surrounding blocking, slicing and writing to \emph{parts} of the matrix
        instead of the whole. This is already prevalent with `dsymm' like BLAS
        routines which only read the lower or upper triangle of a matrix.  This
        idea is inspired by Conor McBride's talk on writing to
        terminals with ``Space Monads''~\cite{space_monads}.

    \item \textbf{Compiling to hardware} is also an option - once we know
        exactly when and where temporaries are required and what can be re-used
        when, we come one (small, but useful) step closer to realising matrix
        expressions directly on hardware.

\end{itemize}