### 2018-05-28 write-up/abstract.tex
Separate out acknowledgements onto its own sheet

% {\normalfont\Huge\sffamily\bfseries Abstract}
% \vspace{24pt}
\begin{abstract}

In this thesis, I argue that linear types are an appropriate, \emph{type-based
formalism} for expressing aliasing, read/write permissions, memory allocation,
re-use and deallocation, first, in the context of the APIs of linear algebra
libraries and then in the context of matrix expression compilation.

I show that framing the problem using linear types can \emph{reduce bugs}
by making precise and explicit, the informal, ad-hoc practices typically
employed by experts and matrix expression compilers \emph{and} automate
checking them.

As evidence for this argument, I show non-trivial, yet readable, linear algebra
programs, that are safe and explicit (with respect to aliasing, read/write
permissions, memory allocation, re-use and deallocation) which (1) are more
memory-efficient than equivalent programs written using high-level
linear algebra libraries and (2) perform just as predictably as equivalent
programs written using low-level linear algebra libraries. I also argue
\emph{the experience} of writing such programs with linear types is
qualitatively better in key respects. In addition to all of this, I show that
it is possible to provide such features \emph{as a library} on top of existing
programming languages and linear algebra libraries.

\end{abstract}

\vspace*{\fill}
\cleardoublepage%

{\normalfont\Huge\sffamily\bfseries Acknowledgements}
\vspace{24pt}

I would like to thank my family and closest friends for supporting me through
my time at Cambridge: you were my strength when I had none. I would also like
to thank my supervisors, Dr\@.~Neelakantan Krishnaswami and Dr\@.~Stephen
Dolan, for teaching me so much this year. I owe a debt of gratitude to both the
2017 Part III Appeals Committee and those who wrote letters in support of my
appeal, for believing I deserved the chance to stay on: never have I valued the
privilege of learning what I enjoy and enjoying what I learn more.  Lastly, I
would like to thank the many staff and research members of the Computer Lab who
have been approachable, patient and attentive during my time here.

\vspace*{\fill}
\cleardoublepage%