### 2018-10-16 write-up/samplepaper/intro.tex
Sketch out structure of paper

\section{Introduction}

\subsection{Contributions}
\begin{itemize}
    \item We have designed the \LANG\ programming language
    \item We illustrate that the design is sensible with many matrix-y examples
    \item We give a soundness proof for \LANG, using a step-indexed logical relation
    \item Incredibly simple type inference algorihtm for polymorphic fractional permissions
        \begin{itemize}
            \item Compare to Bierhof et al's \emph{Fraction Polymorphic Permission Inference},
                which uses a fancy dataflow analysis
            \item We use exactly the same unification algorithm type polymorphism does
        \end{itemize}
    \item We have an implementation - compatible with and usable from existing code!
\end{itemize}