### 2018-10-16 write-up/samplepaper/lang_and_examples.tex

\section{Language Overview and Examples}

\subsection{Short language description}

This should be a high-level description of the language. We should
aim to spend about a page on this section, highlighting the key
properties of \LANG{}:
\begin{itemize}
\item Linearity, obviously
\item Fractional permissions: sharing and unsharing
\item Dynamic errors: matrix dimensions and unsharing (matrix identity checks)
  \begin{itemize}
  \item In the related work, we should point to L3 statically tracking
    pointer identities, but we want to keep the number of type indices
    under control
  \end{itemize}
\end{itemize}

\subsection{Examples}

The more examples, the better. In fact, it's actually impossible to
have too many examples -- it's okay (indeed, desirable) to spend 5-6
pages on examples.

\begin{itemize}

    \item Simple: factorial, shows recursion, !-annotations etc.

    \item Less simple: summing over an array, indexing and safety

    \item Medium: one-dimensioal convolution, permissions

    \item Harder: linear regression, pattern-matching and apparent non-linearity

    \item Harder: L1 norm-minimisation, some frees, re-using memory

    \item Big finish: Kalman filter

\end{itemize}