### 2018-10-21 write-up/paper/paper.tex

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.20 of 2017/10/04
%
\documentclass[runningheads]{llncs}
%
\usepackage{graphicx}
% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
\usepackage{hyperref}
\hypersetup{%
    colorlinks=true,
    linkcolor={black},
    citecolor={green}
}
\renewcommand\UrlFont{\color{blue}\rmfamily}

% Source code highlighting
\usepackage[outputdir=../build]{minted}
  % Magic incantation to stop minted from putting red boxes around shit
  \usepackage{etoolbox}
  \makeatletter
  \AtBeginEnvironment{minted}{\dontdofcolorbox}
  \def\dontdofcolorbox{\renewcommand\fcolorbox[4][]{##4}}
  \makeatother
\RecustomVerbatimEnvironment{Verbatim}{BVerbatim}{}

\newmintinline[highl]{ocaml}{breaklines}

\usepackage{amsmath}

\begin{document}

\title{NumLin: Linear Types for Numerical Libraries}

% \titlerunning{NumLin: Linear Types for Numerical Libraries}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here

\newcommand{\lang}{\textsc{NumLin}}

\author{Dhruv C.~Makwana\inst{1}\orcidID{orcidID} \and
Neelakantan R.~Krishnaswami\inst{2}\orcidID{orcidID}}
%
\authorrunning{D.C. Makwana and N.R. Krishnaswami}

\institute{\email{dcm41@cam.ac.uk}
\url{dhruvmakwana.com} \and
Department of Computer Science, University of Cambridge\\
\email{nk480@cl.cam.ac.uk}}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
Briefly summarize the contents of the paper in 150--250 words.

\keywords{numerical, linear, algebra, types, permissions, OCaml}
\end{abstract}
%
%
%
\input{intro}

\input{lang_and_examples}

\input{formal_system}

\input{implementation}

\input{discussion}

\bibliographystyle{splncs04}

\bibliography{ourbib}

\end{document}