### 2018-10-25 write-up/paper/implementation.tex
Start desugaring

\section{Implementation}\label{sec:implementation}

Talk about how you implemented \lang\ and the general
architecture. Talk about how simple everything is, and also
about how implementing inference for fractions is.

\subsection{Implementation Strategy}

\lang\ transpiles to OCaml and its implementation follows the structure of a
typical domain-specific language (DSL) compiler. Although \lang's current
implementation is not as embedded DSL, its the general design is simple enough
to adapt to being so and also to target other languages.

Alongside the transpiler, a `Read-Check-Translate' loop, benchmarking program
and a test suite are included in the artifacts accompanying this paper.

\begin{enumerate}

    \item \textbf{Parsing}. A generated, LR(1) parser parses a text file into a
        syntax tree. In general, this part will vary for different languages
        and can also be dealt with using combinators or syntax-extensions (the
        EDSL approach) if the host language offers such support.

    \item \textbf{Desugaring}. The syntax tree is then desugared into a
        smaller, more concise, abstract syntax tree. This allows for the type
        checker to be simpler to specify and easier to implement.

    \item \textbf{Matrix Expressions} are also desugared into the abstract
        syntax tree through some simple pattern-matching.

    \item \textbf{Type checking}. The abstract syntax tree is explicitly typed,
        with some inference to make writing typical programs more convenient.

    \item \textbf{Code Generation}. The abstract syntax tree is translated into
        OCaml, with a few `optimisations' to produce more readable code. This
        process is type-preserving: \lang's type system is embedded into
        OCaml's (Figure~\ref{fig:type_grammar}), and so the OCaml type checker
        acts as a sanity check on the generated code.

\end{enumerate}

A very pleasant way to use \lang\ is to have the build system generate code at
\emph{compile-time} and then have the generated code be used by other modules
like normal OCaml functions. This makes it possible and even easy to use \lang\
alongside existing OCaml libraries; in fact, this is exactly how the
benchmarking program and test-suite use code written in \lang.

\subsubsection{Desugaring}\label{subsubsec:desugaring}

Desugaring is conventional.

\begin{align*}
    \mathbf{fun}\ (x : t)\ 'r\ (y : {'r}\ \mathbf{mat}) \rightarrow e
    & \Rightarrow
    \mathbf{fun}\ (x : t) \rightarrow \mathbf{fun}\ {'r} \rightarrow
    \mathbf{fun}\ (y : {'r}\ \mathbf{mat}) \rightarrow e
\\
    x[e]
    & \Rightarrow
    \mathbf{get}\ \_\ x\ (e) \qquad \textrm{(similarly for matrices)}
\\
    x[e_1] := e_2
    & \Rightarrow
    \mathbf{set}\ x\ (e_1)\ (e_2) \qquad \textrm{(similarly for matrices)}
\\
    \mathbf{let}\ !x = e_1\ \mathbf{in}\ e_2
    & \Rightarrow
    \mathbf{let\ Many}\ x = e_1\ \mathbf{in\ }
    \mathbf{let\ Many}\ x = \mathbf{Many}\ (\mathbf{Many}\ x)\ \mathbf{in}\ e_2
\\
    \mathbf{let}\ f\ {\langle args \rangle} = e_1\ \mathbf{in}\ e_2
    & \Rightarrow
    \mathbf{let}\ f = \mathbf{fun}\ {\langle args \rangle} \rightarrow e_1\
    \mathbf{in}\ e_2
\\
    \mathbf{let}\ !f\ {\langle args \rangle} = e_1\ \mathbf{in}\ e_2
    & \Rightarrow
    \mathbf{let\ Many}\ f = \mathbf{Many}\ (\mathbf{fun}\ {\langle args \rangle}
    \rightarrow e_1)\ \mathbf{in}\ e_2
\\
    \mathbf{let\ rec}\ f\ (x : t)\ {\langle args \rangle} : {t'} = e_1\ \mathbf{in}\ e_2
    & \Rightarrow
    \mathbf{let\ Many}\ f = \mathbf{Many}\ (\mathbf{fix}\ (f, x : t,
    \mathbf{fun}\ {\langle args \rangle} \rightarrow e_1 : {t'} ))\ \mathbf{in}\ e_2
\end{align*}

\subsubsection{Matrix Expressions}

Pattern matching!

\subsubsection{Type Checking and Fractional Permission Inference}

Unification!

\subsubsection{Code Generation}

\begin{figure}[tp]
    \centering
    \begin{minipage}{.3\textwidth}
        \centering
        \begin{grammar}
            <f> ::= `'
            \alt <fc>
            \alt `Z'
            \alt `S' <f>

            <t> ::= `'
            \alt `unit'
            \alt `bool'
            \alt `int'
            \alt `elt'
            \alt <f> `arr'
            \alt <f> `mat'
            \alt `!' <t>
            \alt \synt{$\forall$} <fc.> <t>
            \alt <t> \lit{$\otimes$} \synt{$t'$}
            \alt <t> \lit{$\multimap$} \synt{$t'$}
        \end{grammar}
    \end{minipage}
    \begin{minipage}{.3\textwidth}
        \centering
        \begin{minted}[fontsize=\small]{ocaml}
module Arr =
  Owl.Dense.Ndarray.D

type z = Z
type 'a s = Succ

type 'a arr =
  A of Arr.arr
  [@@unboxed]

type 'a mat =
  M of Arr.arr
  [@@unboxed]

type 'a bang =
  Many of 'a
  [@@unboxed]
        \end{minted}
    \end{minipage}
    \begin{minipage}{.3\textwidth}
        \begin{align*}
            [\![ f\!c ]\!] &= \texttt{'fc} \\
            [\![ \textbf{Z} ]\!] &= \texttt{z}\\
            [\![ \textbf{S} \, f ]\!] &= [\![ f ]\!]\, \texttt{s}\\
            [\![ \textbf{unit} ]\!] &= \texttt{unit}\\
            [\![ \textbf{bool} ]\!] &= \texttt{bool}\\
            [\![ \textbf{int} ]\!] &= \texttt{int}\\
            [\![ \textbf{elt} ]\!] &= \texttt{float}\\
            [\![ f\, \textbf{arr} ]\!] &= [\![ f ]\!]\, \texttt{arr}\\
            [\![ f\, \textbf{mat} ]\!] &= [\![ f ]\!]\, \texttt{mat}\\
            [\![ \textbf{!} \, t ]\!] &= [\![ t ]\!]\, \texttt{bang}\\
            [\![ \forall f\!c.\, t ]\!] &= [\![ t ]\!]\\
            [\![ t \otimes t' ]\!] &= [\![ t ]\!] \texttt{*} [\![ t' ]\!]\\
            [\![ t \multimap t' ]\!] &= [\![ t ]\!] \rightarrow [\![ t' ]\!]
        \end{align*}
    \end{minipage}
    \caption{\lang's type grammar (left) and its embedding into OCaml
        (right).}\label{fig:type_grammar}
\end{figure}
A few examples?

\subsection{Performance Metrics}

Here, evaluate the performance of the examples from the second
section.  Compare with your C implementations, and perhaps as well as
the straightforward math transcribed into (Matlab/R/Numpy?).