### 2018-10-28 write-up/paper/intro.tex

\section{Introduction}

\lang\ is a functional programming language designed to express the APIs of
low-level linear algebra libraries (such as BLAS/LAPACK) safely and explicitly.
It does so by combining linear types, fractional permissions, runtime errors
and recursion into a small, easily understandable, yet expressive set of core
constructs. \lang\ allows a novice to understand and work with complicated
linear algebra library APIs; as well as point out subtle aliasing bugs and
reduce memory usage in existing programs. In addition to this, \lang's
implementation supports several syntactic conveniences as well as a
\emph{usable} integration with real OCaml libraries.

\subsection{Contributions}

In this paper
\begin{itemize}
    \item we describe \lang, a linearly typed language for linear algebra programs
    \item we illustrate that \lang's design and features are well-suited to its
        intended domain with progressively sophisticated examples
    \item we prove \lang's soundness, using a step-indexed logical relation
    \item we describe a very simple, unification based type-inference algorithm
        for polymorphic fractional permissions (similar to ones used for
        parametric polymorphism), demonstrating an alternative approach to
        dataflow analysis \cite{bierhoff}
    \item we describe an implementation that is both compatible with and usable
        from existing code.
\end{itemize}