### 2018-11-01 write-up/paper/formal_system.tex
Add interpretation and proof

\section{Formal System}\label{sec:formal_system}

\subsection{Core Type Theory}

The full typing rules are in Appendix \ref{subsec:static_sem}, but the key
ideas are: \textsf{Perm}, \textsf{Type}, $\Theta, \Delta, \Gamma$,
value restriction, lambda, Gen/Spc, and Fix.

\subsection{Dynamic Semantics}\label{subsec:semantics}

and operational semantics here

\subsection{Logical Relation}

Describe the step-indexed logical relation and its main properties

\subsection{Soundness Theorem}

State the fundamental lemma, and sketch the proof a little