### 2018-11-01 write-up/paper/paper.tex

\documentclass[runningheads,a4paper,twoside]{llncs}
\usepackage{hyperref}
\hypersetup{%
    colorlinks=true,
    linkcolor={black},
    citecolor={black},
    urlcolor={blue}
}

% Source code highlighting
\usepackage[outputdir=../build]{minted}
  % Magic incantation to stop minted from putting red boxes around shit
  \usepackage{etoolbox}
  \makeatletter
  \AtBeginEnvironment{minted}{\dontdofcolorbox}
  \def\dontdofcolorbox{\renewcommand\fcolorbox[4][]{##4}}
  \makeatother
\RecustomVerbatimEnvironment{Verbatim}{BVerbatim}{}
% Convenient inline syntax highlighting
\newmintinline[highl]{ocaml}{breaklines}

\usepackage{amsmath,amssymb}

% Ott Rules
\usepackage{supertabular}%
\usepackage[margin=1in]{geometry}%
\usepackage{ifthen}%
\usepackage{alltt}%hack%

\input{../semantics/semantics_def}%
\usepackage{ottlayout}%
\renewcommand{\ottpremise}[1]{\premiseSTY{#1}}%
\renewcommand{\ottusedrule}[1]{\usedruleSTY{#1}}%
\renewcommand{\ottdrule}[4][]{\druleSTY[#1]{#2}{#3}{#4}}%
\renewenvironment{ottdefnblock}[3][]{\defnblockSTY[#1]{#2}{#3}}{\enddefnblockSTY}%

% Proof macros
\newcommand{\den}[3]{ \mathcal{#1}_{#2} [\![ #3 ]\!] }%
\newcommand{\V}[2]{ \den{V}{#1}{#2} }%
\newcommand{\C}[2]{ \den{C}{#1}{#2} }%

\newcommand{\Unit}{\ottkw{unit}}%
\newcommand{\Bang}{\ottkw{!}}
\newcommand{\Bool}{\ottkw{bool}}%
\newcommand{\Int}{\ottkw{int}}%
\newcommand{\Elt}{\ottkw{elt}}%
\newcommand{\Mat}{\ottkw{mat}}%
\newcommand{\Zf}{\ottkw{Z}}%
\newcommand{\Sf}{\ottkw{S}}%
\newcommand{\Many}{\ottkw{Many}}%
\newcommand{\dom}{\mathrm{dom}}%
\newcommand{\empH}{\emptyset}%

\usepackage{pf2}
\beforePfSpace{15pt, 10pt, 10pt, 10pt, 5pt, 2pt}
\afterPfSpace{15pt, 10pt, 10pt, 10pt, 5pt, 2pt}
\interStepSpace{15pt, 10pt, 10pt, 10pt, 5pt, 2pt}
\pflongindent%

% Multi-line table cells
% tex.stackexchange.com/questions/2441/how-to-add-a-forced-line-break-inside-a-table-cell#19678
\newcommand{\specialcell}[3][c]{%
  \begin{array}[#1]{@{}#2@{}}#3\end{array}}

\newcommand{\alsocell}[3][c]{%
  \begin{tabular}[#1]{@{}#2@{}}#3\end{tabular}}

% Figures
\usepackage{graphicx}
\usepackage[dvipsnames]{xcolor}
\usepackage{lscape}
\usepackage{pgfplots}

% PL Stuff
\usepackage[nounderscore]{syntax}
\renewcommand{\syntleft}{\normalfont\itshape}
\renewcommand{\syntright}{}
\renewcommand{\ulitleft}{\normalfont\bf}%\syn@ttspace\frenchspacing}
\renewcommand{\ulitright}{}
\renewcommand{\litleft}{\bgroup\ulitleft}
\renewcommand{\litright}{\ulitright\egroup}

\begin{document}

\title{NumLin: Linear Types for Linear Algebra}

% \titlerunning{NumLin: Linear Types for Linear Algebra}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here

\newcommand{\lang}{\textsc{NumLin}}

\author{Dhruv C.~Makwana\inst{1}\orcidID{orcidID} \and
Neelakantan R.~Krishnaswami\inst{2}\orcidID{orcidID}}
%
\authorrunning{D.C. Makwana and N.R. Krishnaswami}

\institute{\email{dcm41@cam.ac.uk}
\url{dhruvmakwana.com} \and
Department of Computer Science, University of Cambridge\\
\email{nk480@cl.cam.ac.uk}}
%
\maketitle
%
\begin{abstract}
Briefly summarize the contents of the paper in 150--250 words.
%
\keywords{numerical, linear, algebra, types, permissions, OCaml}
\end{abstract}

\input{intro}

\input{lang_and_examples}

\input{formal_system}

\input{implementation}

\input{discussion}

\bibliographystyle{splncs04}

\bibliography{ourbib}

\appendix
\input{appendix}

\end{document}