### 2018-11-12 write-up/paper/appendix.tex
Update Appendix code

\section{\lang\ Specification}

\ottstyledefaults{premiselayout=justify}%
\subsection{Static Semantics}\label{subsec:static_sem}
\ottdefnsTypes%

\subsection{Dynamic Semantics}\label{subsec:dyn_sem}
\ottdefnsOpXXSem%

\input{../semantics/interpretation}

\input{../semantics/lemmas}

\input{../semantics/soundness}

\subsection{Well-formed types}
\ottdefnsWellXXFormed%

\section{\lang\ Grammar}\label{sec:grammar_def}
\ottgrammar%

\section{Desugaring \lang}\label{sec:lang_desugar}

\begin{center}
\[\def\arraystretch{1.3}
    \begin{array}{rcl}
    x[e] &
    \Rightarrow &
    \mathbf{get}\ \_\ x\ (e) \;\qquad \textrm{(similarly for matrices)}
\\
    x[e_1] := e_2 &
    \Rightarrow &
    \mathbf{set}\ x\ (e_1)\ (e_2) \quad \textrm{(similarly for matrices)}
\\
\\
    pat & ::= & ()\ \mid\ x\ \mid\ !x\ \mid\ \mathbf{Many\ } pat\ \mid\ (pat, pat)
\\
    \mathbf{let}\ !x = e_1\ \mathbf{in}\ e_2 &
    \Rightarrow &
    \specialcell[t]{l}{\mathbf{let\ Many}\ x = e_1\ \mathbf{in\ } \\
    \mathbf{let\ Many}\ x = \mathbf{Many}\ (\mathbf{Many}\ x)\ \mathbf{in}\ e_2}
\\
    \mathbf{let\ Many} \langle pat_x \rangle\ = e_1\ \mathbf{in}\ e_2 &
    \Rightarrow &
    \specialcell[t]{l}{%
        \mathbf{let\ Many}\ x = x\ \mathbf{in\ } \\
        \mathbf{let\ } \langle pat_x \rangle\ = x\ \mathbf{in\ } e_2}
\\
    \mathbf{let}\ (\langle pat_a \rangle, \langle pat_b \rangle)\  = e_1\ \mathbf{in}\ e_2 &
    \Rightarrow &
    \specialcell[t]{l}{%
        \mathbf{let\ } (a,b)\ = a\_b\ \mathbf{in\ }
        \ \mathbf{let\ } \langle pat_a \rangle\ = a\ \mathbf{in\ } \\
        \mathbf{let\ } \langle pat_b \rangle\ = b\ \mathbf{in\ } e_2}
\\
    \mathbf{fun}\ (\langle pat_x \rangle : t) \rightarrow e &
    \Rightarrow &
    \mathbf{fun}\ (x : t) \rightarrow \mathbf{let\ } \langle pat_x \rangle = x\ \mathbf{in\ } e
\\
\\
    arg & ::= & \langle pat \rangle : t\ \mid\ {'x} \textrm{ (fractional permission variable)}
\\
    \mathbf{fun}\ \langle arg_{1 .. n} \rangle \rightarrow e &
    \Rightarrow &
    \mathbf{fun}\ \langle arg_1 \rangle \rightarrow ..
    \ \mathbf{fun}\ \langle arg_n \rangle \rightarrow e
\\
    \mathbf{let}\ f\ {\langle arg_{1 .. n} \rangle} = e_1\ \mathbf{in}\ e_2 &
    \Rightarrow &
    \mathbf{let}\ f = \mathbf{fun}\ {\langle arg_{1 .. n} \rangle} \rightarrow e_1\
    \mathbf{in}\ e_2
\\
    \mathbf{let}\ !f\ {\langle arg_{1 .. n} \rangle} = e_1\ \mathbf{in}\ e_2 &
    \Rightarrow &
    \mathbf{let\ Many}\ f = \mathbf{Many}\ (\mathbf{fun}\ {\langle arg_{1 .. n} \rangle}
    \rightarrow e_1)\ \mathbf{in}\ e_2
\\
    \mathrm{fixpoint} & \equiv & \mathbf{fix}\ (f, x : t, \mathbf{fun}
    \ {\langle arg_{1 .. n} \rangle} \rightarrow e_1 : {t'} )
\\
    \mathbf{let\ rec}\ f\ (x : t)\ {\langle arg_{1 .. n} \rangle} : {t'} = e_1\ \mathbf{in}\ e_2 &
    \Rightarrow &
    \mathbf{let}\ f = \mathrm{fixpoint}\ \mathbf{in}\ e_2
\\
    \mathbf{let\ rec}\ !f\ (x : t)\ {\langle arg_{1 .. n} \rangle} : {t'} = e_1\ \mathbf{in}\ e_2 &
    \Rightarrow &
    \mathbf{let\ Many}\ f = \mathbf{Many}\ \mathrm{fixpoint}\ \mathbf{in}\ e_2
    \end{array}
\]
\end{center}

\clearpage
\section{Primitives}\label{subsec:primitives}

\vspace*{\fill}
\begin{center}
\begin{minted}[fontsize=\small]{ocaml}
module Arr = Owl.Dense.Ndarray.D
type z = Z
type 'a s = Succ
type 'a arr = A of Arr.arr [@@unboxed]
type 'a mat = M of Arr.arr [@@unboxed]
type 'a bang = Many of 'a [@@unboxed]
module Prim :
sig
  val extract : 'a bang -> 'a
  (** Boolean *)
  val not_ : bool bang -> bool bang
  (** Arithmetic, many omitted for brevity *)
  val addI : int bang -> int bang -> int bang
  val eqI : int bang -> int bang -> bool bang
  (** Arrays *)
  val set : z arr -> int bang -> float bang -> z arr
  val get : 'a arr -> int bang -> 'a arr * float bang
  val share : 'a arr -> 'a s arr * 'a s arr
  val unshare : 'a s arr -> 'a s arr -> 'a arr
  val free : z arr -> unit
  (** Owl *)
  val array : int bang -> z arr
  val copy : 'a arr -> 'a arr * z arr
  val sin : z arr -> z arr
  val hypot : z arr -> 'a arr -> 'a arr * z arr
  (** Level 1 BLAS *)
  val asum : 'a arr -> 'a arr * float bang
  val axpy : float bang -> 'a arr -> z arr -> 'a arr * z arr
  val dot : 'a arr -> 'b arr -> ('a arr * 'b arr) * float bang
  val scal : float bang -> z arr -> z arr
  val amax : 'a arr -> 'a arr * int bang
  (* Matrix, some omitted for brevity *)
  val matrix : int bang -> int bang -> z mat
  val eye : int bang -> z mat
  val copy_mat : 'a mat -> 'a mat * z mat
  val copy_mat_to : 'a mat -> z mat -> 'a mat * z mat
  val size_mat : 'a mat -> 'a mat * (int bang * int bang)
  val transpose : 'a mat -> 'a mat * z mat
  (* Level 3 BLAS/LAPACK *)
  val gemm : float bang -> ('a mat * bool bang) -> ('b mat * bool bang) ->
             float bang -> z mat -> ('a mat * 'b mat) * z mat
  val symm : bool bang -> float bang -> 'a mat -> 'b mat -> float bang ->
             z mat -> ('a mat * 'b mat) * z mat
  val gesv : z mat -> z mat -> z mat * z mat
  val posv : z mat -> z mat -> z mat * z mat
  val potrs : 'a mat -> z mat -> 'a mat * z mat
  val syrk : bool bang -> float bang -> 'a mat -> float bang -> z mat ->
           'a mat * z mat
end
\end{minted}
\end{center}
\vfill

\clearpage
\section{Kalman Filters from \lang\ and C}

\vspace*{\fill}
\begin{figure}[h]
\begin{center}
    \begin{minted}[fontsize=\footnotesize]{ocaml}
let kalman sigma h mu r_1 data_1 =
  let h, _p_k_n_p_ = Prim.size_mat h in
  let k, n = _p_k_n_p_ in
  let sigma_hT = Prim.matrix n k in
  let (sigma, h), sigma_hT =
    Prim.gemm (Many 1.) (sigma, Many false) (h, Many true) (Many 0.) sigma_hT in
  let (h, sigma_hT), r_2 =
    Prim.gemm (Many 1.) (h, Many false) (sigma_hT, Many false) (Many 1.) r_1 in
  let k_by_k, x = Prim.posv_flip r_2 sigma_hT in
  let (h, mu), data_2 =
    Prim.gemm (Many 1.) (h, Many false) (mu, Many false) (Many (-1.)) data_1 in
  let (x, data_2), new_mu =
    Prim.gemm (Many 1.) (x, Many false) (data_2, Many false) (Many 1.) mu in
  let x_h = Prim.matrix n n in
  let (x, h), x_h =
    Prim.gemm (Many 1.) (x, Many false) (h, Many false) (Many 0.) x_h in
  let () = Prim.free_mat x in
  let sigma, sigma2 = Prim.copy_mat sigma in
  let (sigma, x_h), new_sigma =
    Prim.symm (Many true) (Many (-1.)) sigma x_h (Many 1.) sigma2 in
  let () = Prim.free_mat x_h in
  ((sigma, h), (new_sigma, (new_mu, (k_by_k, data_2))))
    \end{minted}
    \caption{OCaml code for a Kalman filter, generated (at \emph{compile time})
        from the code in Figure~\ref{fig:lang_kalman}, passed through
        \texttt{ocamlformat} for presentation.}\label{fig:ocaml_kalman}
\end{center}
\end{figure}
\vfill

\begin{landscape}
\begin{figure}[p]
    \centering
    \begin{minted}[fontsize=\footnotesize]{c}
static void kalman( const int n,               const int k,                const double *sigma, /* n,n */
                    const double *h, /* k,n */ const double *mu,  /* n,1 */ double *r,           /* k,k */
                    double *data,    /* k,1 */ double **ret_sigma /* n,n */ ) {
    double* n_by_k = (double *) malloc(n * k * sizeof(double));
    cblas_dgemm(CblasRowMajor, CblasNoTrans, CblasTrans, n, k, n, 1., sigma, n, h, n, 0., n_by_k, k);
    cblas_dgemm(CblasRowMajor, CblasNoTrans, CblasNoTrans, k, k, n, 1., h, n, n_by_k, k, 1., r, k);
    LAPACKE_dposv(LAPACK_COL_MAJOR, 'U', k, n, r, k, n_by_k, k);
    cblas_dgemm(CblasRowMajor, CblasNoTrans, CblasNoTrans, k, 1, n, 1., h, n, mu, 1, -1., data, 1);
    cblas_dgemm(CblasRowMajor, CblasNoTrans, CblasNoTrans, n, 1, k, 1., n_by_k, k, data, 1, 1., mu, 1);
    double* n_by_n = (double *) malloc(n * n * sizeof(double));
    cblas_dgemm(CblasRowMajor, CblasNoTrans, CblasNoTrans, n, n, k, 1., n_by_k, k, h, n, 0., n_by_n, n);
    free(n_by_k);
    double* n_by_n2 = (double *) malloc(n * n * sizeof(double));
    cblas_dcopy(n*n, sigma, 1, n_by_n2, 1);
    cblas_dsymm(CblasRowMajor, CblasRight, CblasUpper, n, n, -1., sigma, n, n_by_n, n, 1., n_by_n2, n);
    free(n_by_n);
    *ret_sigma = n_by_n2; }
    \end{minted}
    \caption{CBLAS/LAPACKE implementation of a Kalman filter. I used C instead
        of Fortran because it is what Owl uses under the hood and OCaml FFI
        support for C is better and easier to use than that for Fortran. A distinct
        `measure\_kalman' function that sandwiches a call to this function with
        \texttt{getrusage} is omitted for brevity.}\label{fig:cblas_kalman}

\end{figure}
\end{landscape}