### 2019-01-10 write-up/paper/paper.tex
Update to LIPIcs style

\documentclass[a4paper,UKenglish]{lipics-v2019}
%This is a template for producing LIPIcs articles. 
%See lipics-manual.pdf for further information.
%for A4 paper format use option "a4paper", for US-letter use option "letterpaper"
%for british hyphenation rules use option "UKenglish", for american hyphenation rules use option "USenglish"
%for section-numbered lemmas etc., use "numberwithinsect"
%for enabling cleveref support, use "cleveref"
%for enabling cleveref support, use "autoref"

\bibliographystyle{plainurl}

\title{NumLin: Linear Types for Linear Algebra}

\author{Dhruv C.~Makwana}{Unaffiliated \url{dhruvmakwana.com} }{dcm41@cam.ac.uk}{[orcid]}{}

\author{Neelakantan R.~Krishnaswami}{Department of Computer Science and Technology, University of Cambridge, United Kingdom}{nk480@cl.cam.ac.uk}{[orcid]}{}

\authorrunning{D.\,C. Makwana and N.\,R. Krishnaswami}

\Copyright{Dhruv C. Makwana and Neelakantan R. Krishnaswami}% LIPIcs license is "CC-BY";  http://creativecommons.org/licenses/by/3.0/

\ccsdesc[300]{Theory of computation~Program specifications}

\keywords{numerical, linear, algebra, types, permissions, OCaml}

\supplement{\url{www.github.com/dc-mak/lt4la}}

%\nolinenumbers %uncomment to disable line numbering

%\hideLIPIcs  %uncomment to remove references to LIPIcs series (logo, DOI, ...), e.g. when preparing a pre-final version to be uploaded to arXiv or another public repository

%Editor-only macros:: begin (do not touch as author)%
\EventEditors{John Q. Open and Joan R. Access}
\EventNoEds{2}
\EventLongTitle{42nd Conference on Very Important Topics (CVIT 2016)}
\EventShortTitle{CVIT 2016}
\EventAcronym{CVIT}
\EventYear{2016}
\EventDate{December 24--27, 2016}
\EventLocation{Little Whinging, United Kingdom}
\EventLogo{}
\SeriesVolume{42}
\ArticleNo{23}

% Source code highlighting
\usepackage[outputdir=../build]{minted}
  % Magic incantation to stop minted from putting red boxes around shit
  \usepackage{etoolbox}
  \makeatletter
  \AtBeginEnvironment{minted}{\dontdofcolorbox}
  \def\dontdofcolorbox{\renewcommand\fcolorbox[4][]{##4}}
  \makeatother
\RecustomVerbatimEnvironment{Verbatim}{BVerbatim}{}
% Convenient inline syntax highlighting
\newmintinline[highl]{ocaml}{breaklines}

\usepackage{amsmath,amssymb}

% Ott Rules
\usepackage{supertabular}%
% \usepackage{geometry}%
\usepackage{ifthen}%
\usepackage{alltt}%hack%

\input{../semantics/semantics_def}%
\usepackage{ottlayout}%
\renewcommand{\ottpremise}[1]{\premiseSTY{#1}}%
\renewcommand{\ottusedrule}[1]{\usedruleSTY{#1}}%
\renewcommand{\ottdrule}[4][]{\druleSTY[#1]{#2}{#3}{#4}}%
\renewenvironment{ottdefnblock}[3][]{\defnblockSTY[#1]{#2}{#3}}{\enddefnblockSTY}%

% Proof macros
\newcommand{\den}[3]{ \mathcal{#1}_{#2} [\![ #3 ]\!] }%
\newcommand{\V}[2]{ \den{V}{#1}{#2} }%
\newcommand{\C}[2]{ \den{C}{#1}{#2} }%

\newcommand{\Unit}{\ottkw{unit}}%
\newcommand{\Bang}{\ottkw{!}}
\newcommand{\Bool}{\ottkw{bool}}%
\newcommand{\Int}{\ottkw{int}}%
\newcommand{\Elt}{\ottkw{elt}}%
\newcommand{\Mat}{\ottkw{mat}}%
\newcommand{\Zf}{\ottkw{Z}}%
\newcommand{\Sf}{\ottkw{S}}%
\newcommand{\Many}{\ottkw{Many}}%
\newcommand{\dom}{\mathrm{dom}}%
\newcommand{\empH}{\emptyset}%

\usepackage{pf2}
\beforePfSpace{15pt, 10pt, 10pt, 10pt, 5pt, 2pt}
\afterPfSpace{15pt, 10pt, 10pt, 10pt, 5pt, 2pt}
\interStepSpace{15pt, 10pt, 10pt, 10pt, 5pt, 2pt}
\pflongindent%

% Multi-line table cells
% tex.stackexchange.com/questions/2441/how-to-add-a-forced-line-break-inside-a-table-cell#19678
\newcommand{\specialcell}[3][c]{%
  \begin{array}[#1]{@{}#2@{}}#3\end{array}}

\newcommand{\alsocell}[3][c]{%
  \begin{tabular}[#1]{@{}#2@{}}#3\end{tabular}}

% Figures
\usepackage{graphicx}
\usepackage[dvipsnames]{xcolor}
\usepackage{lscape}
\usepackage{pgfplots}

% PL Stuff
\usepackage[nounderscore]{syntax}
\renewcommand{\syntleft}{\normalfont\itshape}
\renewcommand{\syntright}{}
\renewcommand{\ulitleft}{\normalfont\bf}%\syn@ttspace\frenchspacing}
\renewcommand{\ulitright}{}
\renewcommand{\litleft}{\bgroup\ulitleft}
\renewcommand{\litright}{\ulitright\egroup}

% NumLin
\newcommand{\lang}{\textsc{NumLin}}

\begin{document}

\begin{abstract}
    We present \lang, a functional programming language designed to express
    the APIs of low-level linear algebra libraries (such as BLAS/LAPACK) safely
    and explicitly, through a brief description of its key features and several
    illustrative examples. We show that \lang's type system is sound and that
    its implementation improves upon na{\"i}ve implementations of linear
    algebra programs, almost towards C-levels of performance. Lastly, we contrast
    it to other recent developments in linear types and show that using linear
    types and fractional permissions to express the APIs of low-level linear
    algebra libraries is a simple and effective idea.
\end{abstract}

\input{intro}

\input{lang_and_examples}

\input{formal_system}

\input{implementation}

\input{discussion}

\clearpage
\bibliography{ourbib}

\clearpage
\appendix
\input{appendix}

\end{document}